\chapterornament
\section*{Page 1}

\lettrine{N}{ow in those days, a decree went out from Caesar Augustus that all the world should be enrolled.}

This was the first enrollment made when Quirinius was governor of Syria.

All went to enroll themselves, everyone to his own city.

Joseph also went up from Galilee, out of the city of Nazareth, into Judea, to David’s city, which is called Bethlehem, because he was of the house and family of David;

to enroll himself with Mary, who was pledged to be married to him as wife, being pregnant.

While they were there, the day had come for her to give birth.

She gave birth to her firstborn son. She wrapped him in bands of cloth, and laid him in a feeding trough, because there was no room for them in the inn.

When eight days were fulfilled for the circumcision of the child, his name was called Jesus, which was given by the angel before he was conceived in the womb.

When they had accomplished all things that were according to the law of the Lord, they returned into Galilee, to their own city, Nazareth.

The child was growing, and was becoming strong in spirit, being filled with wisdom, and the grace of God was upon him.

When he was twelve years old, they went up to Jerusalem according to the custom of the feast,

and when they had fulfilled the days, as they were returning, the boy Jesus stayed behind in Jerusalem. Joseph and his mother didn’t know it,

but supposing him to be in the company, they went a day’s journey, and they looked for him among their relatives and acquaintances.

When they didn’t find him, they returned to Jerusalem, looking for him.

After three days they found him in the temple, sitting in the middle of the teachers, both listening to them, and asking them questions.

All who heard him were amazed at his understanding and his answers.

When they saw him, they were astonished, and his mother said to him, “Son, why have you treated us this way? Behold, your father and I were anxiously looking for you.”

And he went down with them, and came to Nazareth. He was subject to them, and his mother kept all these sayings in her heart.

And Jesus increased in wisdom and stature, and in favor with God and men.

Now in the fifteenth year of the reign of Tiberius Caesar, Pontius Pilate being governor of Judea, and Herod being tetrarch of Galilee, and his brother Philip tetrarch of the region of Ituraea and Trachonitis, and Lysanias tetrarch of Abilene,



\chapterornament
\section*{Page 2}

\lettrine{i}{n the high priesthood of Annas and Caiaphas, the word of God came to John, the son of Zacharias, in the wilderness.}

John came baptizing in the wilderness and preaching the baptism of repentance for forgiveness of sins.

Now John himself wore clothing made of camel’s hair, with a leather belt around his waist. His food was locusts and wild honey.

Then people from Jerusalem, all of Judea, and all the region around the Jordan went out to him.

They were baptized by him in the Jordan, confessing their sins.

Then Jesus came from Galilee to the Jordan to John, to be baptized by him.

Jesus himself, when he began to teach, was about thirty years old, being the son (as was supposed) of Joseph, the son of Heli,

After this, he went down to Capernaum, he, and his mother, his brothers, and his disciples; and they stayed there a few days.

The Passover of the Jews was at hand, and Jesus went up to Jerusalem.

He found in the temple those who sold oxen, sheep, and doves, and the changers of money sitting.

He made a whip of cords, and threw all out of the temple, both the sheep and the oxen; and he poured out the changers’ money and overthrew their tables.

To those who sold the doves, he said, “Take these things out of here! Don’t make my Father’s house a marketplace!”

After these things, Jesus came with his disciples into the land of Judea. He stayed there with them and baptized.

Now when Jesus heard that John was delivered up, he withdrew into Galilee.

For Herod himself had sent out and arrested John, and bound him in prison for the sake of Herodias, his brother Philip’s wife, for he had married her.

For John said to Herod, “It is not lawful for you to have your brother’s wife.”

Herodias set herself against him, and desired to kill him, but she couldn’t,

for Herod feared John, knowing that he was a righteous and holy man, and kept him safe. When he heard him, he did many things, and he heard him gladly.

Then a convenient day came, that Herod on his birthday made a supper for his nobles, the high officers, and the chief men of Galilee.

When the daughter of Herodias herself came in and danced, she pleased Herod and those sitting with him. The king said to the young lady, “Ask me whatever you want, and I will give it to you.”



\chapterornament
\section*{Page 3}

\lettrine{H}{e swore to her, “Whatever you shall ask of me, I will give you, up to half of my kingdom.”}

She went out, and said to her mother, “What shall I ask?” She said, “The head of John the Baptizer.”

She came in immediately with haste to the king, and asked, “I want you to give me right now the head of John the Baptizer on a platter.”

The king was exceedingly sorry, but for the sake of his oaths, and of his dinner guests, he didn’t wish to refuse her.

Immediately the king sent out a soldier of his guard, and commanded to bring John’s head, and he went and beheaded him in the prison,

and brought his head on a platter, and gave it to the young lady; and the young lady gave it to her mother.

They went into Capernaum, and immediately on the Sabbath day he entered into the synagogue and taught.

They were astonished at his teaching, for he taught them as having authority, and not as the scribes.

At that time, Jesus went on the Sabbath day through the grain fields. His disciples were hungry and began to pluck heads of grain and to eat.

But the Pharisees, when they saw it, said to him, “Behold, your disciples do what is not lawful to do on the Sabbath.”

But he said to them, “Haven’t you read what David did, when he was hungry, and those who were with him;

how he entered into God’s house, and ate the show bread, which was not lawful for him to eat, neither for those who were with him, but only for the priests?

Or have you not read in the law, that on the Sabbath day, the priests in the temple profane the Sabbath, and are guiltless?

She went out, and said to her mother, “What shall I ask?” She said, “The head of John the Baptizer.”

She came in immediately with haste to the king, and asked, “I want you to give me right now the head of John the Baptizer on a platter.”

The king was exceedingly sorry, but for the sake of his oaths, and of his dinner guests, he didn’t wish to refuse her.

Immediately the king sent out a soldier of his guard, and commanded to bring John’s head, and he went and beheaded him in the prison,

and brought his head on a platter, and gave it to the young lady; and the young lady gave it to her mother.

They went into Capernaum, and immediately on the Sabbath day he entered into the synagogue and taught.

They were astonished at his teaching, for he taught them as having authority, and not as the scribes.



\chapterornament
\section*{Page 4}

\lettrine{A}{t that time, Jesus went on the Sabbath day through the grain fields. His disciples were hungry and began to pluck heads of grain and to eat.}

But the Pharisees, when they saw it, said to him, “Behold, your disciples do what is not lawful to do on the Sabbath.”

But he said to them, “Haven’t you read what David did, when he was hungry, and those who were with him;

how he entered into God’s house, and ate the show bread, which was not lawful for him to eat, neither for those who were with him, but only for the priests?

Or have you not read in the law, that on the Sabbath day, the priests in the temple profane the Sabbath, and are guiltless?

Seeing the multitudes, he went up onto the mountain. When he had sat down, his disciples came to him.

He opened his mouth and taught them, saying,

“Blessed are the poor in spirit, for theirs is the Kingdom of Heaven.

Blessed are those who mourn, for they shall be comforted.

Blessed are the gentle, for they shall inherit the earth.

Blessed are those who hunger and thirst after righteousness, for they shall be filled.

Blessed are the merciful, for they shall obtain mercy.

Blessed are the pure in heart, for they shall see God.

Blessed are the peacemakers, for they shall be called children of God.

Blessed are those who have been persecuted for righteousness’ sake, for theirs is the Kingdom of Heaven.

“Blessed are you when people reproach you, persecute you, and say all kinds of evil against you falsely, for my sake.

Rejoice, and be exceedingly glad, for great is your reward in heaven. For that is how they persecuted the prophets who were before you.

“But woe to you who are rich! For you have received your consolation.

Woe to you, you who are full now, for you will be hungry. Woe to you who laugh now, for you will mourn and weep.

Woe, when men speak well of you, for their fathers did the same thing to the false prophets.



\chapterornament
\section*{Page 5}

\lettrine{“}{You are the salt of the earth, but if the salt has lost its flavor, with what will it be salted? It is then good for nothing, but to be cast out and trodden under the feet of men.}

You are the light of the world. A city located on a hill can’t be hidden.

Neither do you light a lamp, and put it under a measuring basket, but on a stand; and it shines to all who are in the house.

Even so, let your light shine before men; that they may see your good works, and glorify your Father who is in heaven.

“Don’t think that I came to destroy the law or the prophets. I didn’t come to destroy, but to fulfill.

For most certainly, I tell you, until heaven and earth pass away, not even one smallest letter or one tiny pen stroke shall in any way pass away from the law, until all things are accomplished.

Whoever, therefore, shall break one of these least commandments, and teach others to do so, shall be called least in the Kingdom of Heaven; but whoever shall do and teach them shall be called great in the Kingdom of Heaven.

For I tell you that unless your righteousness exceeds that of the scribes and Pharisees, there is no way you will enter into the Kingdom of Heaven.

“You have heard that it was said to the ancient ones, ‘You shall not murder;’ and ‘Whoever murders will be in danger of the judgment.’

But I tell you that everyone who is angry with his brother without a cause   will be in danger of the judgment. Whoever says to his brother, ‘Raca!’  will be in danger of the council. Whoever says, ‘You fool!’ will be in danger of the fire of Gehenna.

“If therefore you are offering your gift at the altar, and there remember that your brother has anything against you,

leave your gift there before the altar, and go your way. First be reconciled to your brother, and then come and offer your gift.

Agree with your adversary quickly, while you are with him on the way; lest perhaps the prosecutor deliver you to the judge, and the judge deliver you to the officer, and you be cast into prison.

Most certainly I tell you, you shall by no means get out of there, until you have paid the last penny.

“You have heard that it was said,  ‘You shall not commit adultery;’

but I tell you that everyone who gazes at a woman to lust after her has committed adultery with her already in his heart.

If your right eye causes you to stumble, pluck it out and throw it away from you. For it is more profitable for you that one of your members should perish, than for your whole body to be cast into Gehenna.

If your right hand causes you to stumble, cut it off, and throw it away from you. For it is more profitable for you that one of your members should perish, than for your whole body to be cast into Gehenna.

“It was also said, ‘Whoever shall put away his wife, let him give her a writing of divorce,’

but I tell you that whoever puts away his wife, except for the cause of sexual immorality, makes her an adulteress; and whoever marries her when she is put away commits adultery.



\chapterornament
\section*{Page 6}

\lettrine{“}{Again you have heard that it was said to the ancient ones, ‘You shall not make false vows, but shall perform to the Lord your vows,’}

but I tell you, don’t swear at all: neither by heaven, for it is the throne of God;

nor by the earth, for it is the footstool of his feet; nor by Jerusalem, for it is the city of the great King.

Neither shall you swear by your head, for you can’t make one hair white or black.

But let your ‘Yes’ be ‘Yes’ and your ‘No’ be ‘No.’ Whatever is more than these is of the evil one.

“You have heard that it was said, ‘An eye for an eye, and a tooth for a tooth.’

But I tell you, don’t resist him who is evil; but whoever strikes you on your right cheek, turn to him the other also.

If anyone sues you to take away your coat, let him have your cloak also.

Whoever compels you to go one mile, go with him two.

Give to him who asks you, and don’t turn away him who desires to borrow from you.

“You have heard that it was said, ‘You shall love your neighbor  and hate your enemy.’

But I tell you, love your enemies, bless those who curse you, do good to those who hate you, and pray for those who mistreat you and persecute you,

that you may be children of your Father who is in heaven. For he makes his sun to rise on the evil and the good, and sends rain on the just and the unjust.

For if you love those who love you, what reward do you have? Don’t even the tax collectors do the same?

If you only greet your friends, what more do you do than others? Don’t even the tax collectors do the same?

If you lend to those from whom you hope to receive, what credit is that to you? Even sinners lend to sinners, to receive back as much.

But love your enemies, and do good, and lend, expecting nothing back; and your reward will be great, and you will be children of the Most High; for he is kind toward the unthankful and evil.

“Therefore be merciful, even as your Father is also merciful.

“Be careful that you don’t do your charitable giving before men, to be seen by them, or else you have no reward from your Father who is in heaven.

Therefore when you do merciful deeds, don’t sound a trumpet before yourself, as the hypocrites do in the synagogues and in the streets, that they may get glory from men. Most certainly I tell you, they have received their reward.



\chapterornament
\section*{Page 7}

\lettrine{B}{ut when you do merciful deeds, don’t let your left hand know what your right hand does,}

so that your merciful deeds may be in secret, then your Father who sees in secret will reward you openly.

“When you pray, you shall not be as the hypocrites, for they love to stand and pray in the synagogues and in the corners of the streets, that they may be seen by men. Most certainly, I tell you, they have received their reward.

But you, when you pray, enter into your inner room, and having shut your door, pray to your Father who is in secret, and your Father who sees in secret will reward you openly.

In praying, don’t use vain repetitions, as the Gentiles do; for they think that they will be heard for their much speaking.

Therefore don’t be like them, for your Father knows what things you need, before you ask him.

Pray like this: ‘Our Father in heaven, may your name be kept holy.

Let your Kingdom come. Let your will be done on earth as it is in heaven.

Give us today our daily bread.

Forgive us our debts, as we also forgive our debtors.

Bring us not into temptation, but deliver us from the evil one. For yours is the Kingdom, the power, and the glory forever. Amen.’

“For if you forgive men their trespasses, your heavenly Father will also forgive you.

But if you don’t forgive men their trespasses, neither will your Father forgive your trespasses.

“Moreover when you fast, don’t be like the hypocrites, with sad faces. For they disfigure their faces, that they may be seen by men to be fasting. Most certainly I tell you, they have received their reward.

But you, when you fast, anoint your head, and wash your face;

so that you are not seen by men to be fasting, but by your Father who is in secret, and your Father, who sees in secret, will reward you.

“Don’t lay up treasures for yourselves on the earth, where moth and rust consume, and where thieves break through and steal;

but lay up for yourselves treasures in heaven, where neither moth nor rust consume, and where thieves don’t break through and steal;

for where your treasure is, there your heart will be also.

“The lamp of the body is the eye. If therefore your eye is sound, your whole body will be full of light.



\chapterornament
\section*{Page 8}

\lettrine{B}{ut if your eye is evil, your whole body will be full of darkness. If therefore the light that is in you is darkness, how great is the darkness!}

“No one can serve two masters, for either he will hate the one and love the other; or else he will be devoted to one and despise the other. You can’t serve both God and Mammon.

Therefore I tell you, don’t be anxious for your life: what you will eat, or what you will drink; nor yet for your body, what you will wear. Isn’t life more than food, and the body more than clothing?

See the birds of the sky, that they don’t sow, neither do they reap, nor gather into barns. Your heavenly Father feeds them. Aren’t you of much more value than they?

“Which of you, by being anxious, can add one moment to his lifespan?

Why are you anxious about clothing? Consider the lilies of the field, how they grow. They don’t toil, neither do they spin,

yet I tell you that even Solomon in all his glory was not dressed like one of these.

But if God so clothes the grass of the field, which today exists, and tomorrow is thrown into the oven, won’t he much more clothe you, you of little faith?

“Therefore don’t be anxious, saying, ‘What will we eat?’, ‘What will we drink?’ or, ‘With what will we be clothed?’

For the Gentiles seek after all these things; for your heavenly Father knows that you need all these things.

But seek first God’s Kingdom, and his righteousness; and all these things will be given to you as well.

Therefore don’t be anxious for tomorrow, for tomorrow will be anxious for itself. Each day’s own evil is sufficient.

“Don’t judge, so that you won’t be judged.

For with whatever judgment you judge, you will be judged; and with whatever measure you measure, it will be measured to you.

“Give, and it will be given to you: good measure, pressed down, shaken together, and running over, will be given to you. For with the same measure you measure it will be measured back to you.”

Why do you see the speck that is in your brother’s eye, but don’t consider the beam that is in your own eye?

Or how will you tell your brother, ‘Let me remove the speck from your eye;’ and behold, the beam is in your own eye?

You hypocrite! First remove the beam out of your own eye, and then you can see clearly to remove the speck out of your brother’s eye.

“Don’t give that which is holy to the dogs, neither throw your pearls before the pigs, lest perhaps they trample them under their feet, and turn and tear you to pieces.

“Ask, and it will be given you. Seek, and you will find. Knock, and it will be opened for you.



\chapterornament
\section*{Page 9}

\lettrine{F}{or everyone who asks receives. He who seeks finds. To him who knocks it will be opened.}

Or who is there among you, who, if his son asks him for bread, will give him a stone?

Or if he asks for a fish, who will give him a serpent?

If you then, being evil, know how to give good gifts to your children, how much more will your Father who is in heaven give good things to those who ask him!

Therefore whatever you desire for men to do to you, you shall also do to them; for this is the law and the prophets.

“Enter in by the narrow gate; for wide is the gate and broad is the way that leads to destruction, and many are those who enter in by it.

How narrow is the gate, and restricted is the way that leads to life! Few are those who find it.

“Beware of false prophets, who come to you in sheep’s clothing, but inwardly are ravening wolves.

By their fruits you will know them. Do you gather grapes from thorns, or figs from thistles?

Even so, every good tree produces good fruit; but the corrupt tree produces evil fruit.

A good tree can’t produce evil fruit, neither can a corrupt tree produce good fruit.

Every tree that doesn’t grow good fruit is cut down, and thrown into the fire.

Therefore by their fruits you will know them.

The good man out of his good treasure brings out good things, and the evil man out of his evil treasure brings out evil things.

I tell you that every idle word that men speak, they will give account of it in the day of judgment.

For by your words you will be justified, and by your words you will be condemned.”

“Everyone therefore who hears these words of mine, and does them, I will liken him to a wise man, who built his house on a rock.

The rain came down, the floods came, and the winds blew, and beat on that house; and it didn’t fall, for it was founded on the rock.

Everyone who hears these words of mine, and doesn’t do them will be like a foolish man, who built his house on the sand.

The rain came down, the floods came, and the winds blew, and beat on that house; and it fell—and great was its fall.”



\chapterornament
\section*{Page 10}

\lettrine{W}{hen Jesus had finished saying these things, the multitudes were astonished at his teaching,}

for he taught them with authority, and not like the scribes.

When he came down from the mountain, great multitudes followed him.

He marveled because of their unbelief. He went around the villages teaching.

“Come to me, all you who labor and are heavily burdened, and I will give you rest.

Take my yoke upon you, and learn from me, for I am gentle and humble in heart; and you will find rest for your souls.

For my yoke is easy, and my burden is light.”

One of the Pharisees invited him to eat with him. He entered into the Pharisee’s house, and sat at the table.

Behold, a woman in the city who was a sinner, when she knew that he was reclining in the Pharisee’s house, brought an alabaster jar of ointment.

Standing behind at his feet weeping, she began to wet his feet with her tears, and she wiped them with the hair of her head, kissed his feet, and anointed them with the ointment.

Now when the Pharisee who had invited him saw it, he said to himself, “This man, if he were a prophet, would have perceived who and what kind of woman this is who touches him, that she is a sinner.”

Jesus answered him, “Simon, I have something to tell you.” He said, “Teacher, say on.”

“A certain lender had two debtors. The one owed five hundred denarii, and the other fifty.

When they couldn’t pay, he forgave them both. Which of them therefore will love him most?”

Simon answered, “He, I suppose, to whom he forgave the most.” He said to him, “You have judged correctly.”

Turning to the woman, he said to Simon, “Do you see this woman? I entered into your house, and you gave me no water for my feet, but she has wet my feet with her tears, and wiped them with the hair of her head.

You gave me no kiss, but she, since the time I came in, has not ceased to kiss my feet.

You didn’t anoint my head with oil, but she has anointed my feet with ointment.

His mother and his brothers came, and standing outside, they sent to him, calling him.

A multitude was sitting around him, and they told him, “Behold, your mother, your brothers, and your sisters are outside looking for you.”



\chapterornament
\section*{Page 11}

\lettrine{H}{e answered them, “Who are my mother and my brothers?”}

Looking around at those who sat around him, he said, “Behold, my mother and my brothers!

For whoever does the will of God is my brother, my sister, and mother.”

Meanwhile, when a multitude of many thousands had gathered together, so much so that they trampled on each other, he began to tell his disciples first of all, “Beware of the yeast of the Pharisees, which is hypocrisy.

But there is nothing covered up that will not be revealed, nor hidden that will not be known.

Therefore whatever you have said in the darkness will be heard in the light. What you have spoken in the ear in the inner rooms will be proclaimed on the housetops.

“I tell you, my friends, don’t be afraid of those who kill the body, and after that have no more that they can do.

But I will warn you whom you should fear. Fear him who after he has killed, has power to cast into Gehenna. Yes, I tell you, fear him.

“Aren’t five sparrows sold for two assaria coins? Not one of them is forgotten by God.

But the very hairs of your head are all counted. Therefore don’t be afraid. You are of more value than many sparrows.

One of the multitude said to him, “Teacher, tell my brother to divide the inheritance with me.”

But he said to him, “Man, who made me a judge or an arbitrator over you?”

He said to them, “Beware! Keep yourselves from covetousness, for a man’s life doesn’t consist of the abundance of the things which he possesses.”

He spoke a parable to them, saying, “The ground of a certain rich man produced abundantly.

He reasoned within himself, saying, ‘What will I do, because I don’t have room to store my crops?’

He said, ‘This is what I will do. I will pull down my barns, build bigger ones, and there I will store all my grain and my goods.

I will tell my soul, “Soul, you have many goods laid up for many years. Take your ease, eat, drink, and be merry.” ’

“But God said to him, ‘You foolish one, tonight your soul is required of you. The things which you have prepared—whose will they be?’

So is he who lays up treasure for himself, and is not rich toward God.”

He said to his disciples, “Therefore I tell you, don’t be anxious for your life, what you will eat, nor yet for your body, what you will wear.



\chapterornament
\section*{Page 12}

\lettrine{L}{ife is more than food, and the body is more than clothing.}

Consider the ravens: they don’t sow, they don’t reap, they have no warehouse or barn, and God feeds them. How much more valuable are you than birds!

Which of you by being anxious can add a cubit to his height?

If then you aren’t able to do even the least things, why are you anxious about the rest?

Consider the lilies, how they grow. They don’t toil, neither do they spin; yet I tell you, even Solomon in all his glory was not arrayed like one of these.

But if this is how God clothes the grass in the field, which today exists, and tomorrow is cast into the oven, how much more will he clothe you, O you of little faith?

Don’t seek what you will eat or what you will drink; neither be anxious.

For the nations of the world seek after all of these things, but your Father knows that you need these things.

But seek God’s Kingdom, and all these things will be added to you.

Don’t be afraid, little flock, for it is your Father’s good pleasure to give you the Kingdom.

Sell that which you have, and give gifts to the needy. Make for yourselves purses which don’t grow old, a treasure in the heavens that doesn’t fail, where no thief approaches, neither moth destroys.

For where your treasure is, there will your heart be also.

“Let your waist be dressed and your lamps burning.

Be like men watching for their lord, when he returns from the wedding feast; that when he comes and knocks, they may immediately open to him.

Blessed are those servants, whom the lord will find watching when he comes. Most certainly I tell you that he will dress himself, make them recline, and will come and serve them.

They will be blessed if he comes in the second or third watch, and finds them so.

But know this, that if the master of the house had known in what hour the thief was coming, he would have watched, and not allowed his house to be broken into.

Therefore be ready also, for the Son of Man is coming in an hour that you don’t expect him.”

Peter said to him, “Lord, are you telling this parable to us, or to everybody?”

The Lord said, “Who then is the faithful and wise steward, whom his lord will set over his household, to give them their portion of food at the right times?



\chapterornament
\section*{Page 13}

\lettrine{B}{lessed is that servant whom his lord will find doing so when he comes.}

Truly I tell you, that he will set him over all that he has.

But if that servant says in his heart, ‘My lord delays his coming,’ and begins to beat the menservants and the maidservants, and to eat and drink, and to be drunken,

then the lord of that servant will come in a day when he isn’t expecting him, and in an hour that he doesn’t know, and will cut him in two, and place his portion with the unfaithful.

That servant, who knew his lord’s will, and didn’t prepare, nor do what he wanted, will be beaten with many stripes,

but he who didn’t know, and did things worthy of stripes, will be beaten with few stripes. To whomever much is given, of him will much be required; and to whom much was entrusted, of him more will be asked.

He said to the multitudes also, “When you see a cloud rising from the west, immediately you say, ‘A shower is coming,’ and so it happens.

When a south wind blows, you say, ‘There will be a scorching heat,’ and it happens.

You hypocrites! You know how to interpret the appearance of the earth and the sky, but how is it that you don’t interpret this time?

Why don’t you judge for yourselves what is right?

For when you are going with your adversary before the magistrate, try diligently on the way to be released from him, lest perhaps he drag you to the judge, and the judge deliver you to the officer, and the officer throw you into prison.

I tell you, you will by no means get out of there, until you have paid the very last penny.”

Now there were some present at the same time who told him about the Galileans, whose blood Pilate had mixed with their sacrifices.

Jesus answered them, “Do you think that these Galileans were worse sinners than all the other Galileans, because they suffered such things?

I tell you, no, but unless you repent, you will all perish in the same way.

Or those eighteen, on whom the tower in Siloam fell and killed them; do you think that they were worse offenders than all the men who dwell in Jerusalem?

I tell you, no, but, unless you repent, you will all perish in the same way.”

He spoke this parable. “A certain man had a fig tree planted in his vineyard, and he came seeking fruit on it, and found none.

He said to the vine dresser, ‘Behold, these three years I have come looking for fruit on this fig tree, and found none. Cut it down. Why does it waste the soil?’

He answered, ‘Lord, leave it alone this year also, until I dig around it and fertilize it.



\chapterornament
\section*{Page 14}

\lettrine{I}{f it bears fruit, fine; but if not, after that, you can cut it down.’ ”}

Now as he spoke, a certain Pharisee asked him to dine with him. He went in, and sat at the table.

When the Pharisee saw it, he marveled that he had not first washed himself before dinner.

The Lord said to him, “Now you Pharisees cleanse the outside of the cup and of the platter, but your inward part is full of extortion and wickedness.

You foolish ones, didn’t he who made the outside make the inside also?

But give for gifts to the needy those things which are within, and behold, all things will be clean to you.

But woe to you Pharisees! For you tithe mint and rue and every herb, but you bypass justice and God’s love. You ought to have done these, and not to have left the other undone.

Woe to you Pharisees! For you love the best seats in the synagogues, and the greetings in the marketplaces.

Woe to you, scribes and Pharisees, hypocrites! For you are like hidden graves, and the men who walk over them don’t know it.”

One of the lawyers answered him, “Teacher, in saying this you insult us also.”

He said, “Woe to you lawyers also! For you load men with burdens that are difficult to carry, and you yourselves won’t even lift one finger to help carry those burdens.

Woe to you lawyers! For you took away the key of knowledge. You didn’t enter in yourselves, and those who were entering in, you hindered.”

As he said these things to them, the scribes and the Pharisees began to be terribly angry, and to draw many things out of him;

lying in wait for him, and seeking to catch him in something he might say, that they might accuse him.

On that day Jesus went out of the house, and sat by the seaside.

Great multitudes gathered to him, so that he entered into a boat, and sat, and all the multitude stood on the beach.

He spoke to them many things in parables, saying, “Behold, a farmer went out to sow.

As he sowed, some seeds fell by the roadside, and the birds came and devoured them.

Others fell on rocky ground, where they didn’t have much soil, and immediately they sprang up, because they had no depth of earth.

When the sun had risen, they were scorched. Because they had no root, they withered away.



\chapterornament
\section*{Page 15}

\lettrine{O}{thers fell among thorns. The thorns grew up and choked them.}

Others fell on good soil, and yielded fruit: some one hundred times as much, some sixty, and some thirty.

He who has ears to hear, let him hear.”

When he was alone, those who were around him with the twelve asked him about the parables.

“Hear, then, the parable of the farmer.

When anyone hears the word of the Kingdom, and doesn’t understand it, the evil one comes, and snatches away that which has been sown in his heart. This is what was sown by the roadside.

What was sown on the rocky places, this is he who hears the word, and immediately with joy receives it;

yet he has no root in himself, but endures for a while. When oppression or persecution arises because of the word, immediately he stumbles.

What was sown among the thorns, this is he who hears the word, but the cares of this age and the deceitfulness of riches choke the word, and he becomes unfruitful.

What was sown on the good ground, this is he who hears the word, and understands it, who most certainly bears fruit, and produces, some one hundred times as much, some sixty, and some thirty.”

He said to them, “Is the lamp brought to be put under a basket  or under a bed? Isn’t it put on a stand?

For there is nothing hidden, except that it should be made known; neither was anything made secret, but that it should come to light.

If any man has ears to hear, let him hear.”

He set another parable before them, saying, “The Kingdom of Heaven is like a man who sowed good seed in his field,

but while people slept, his enemy came and sowed darnel weeds also among the wheat, and went away.

But when the blade sprang up and produced fruit, then the darnel weeds appeared also.

The servants of the householder came and said to him, ‘Sir, didn’t you sow good seed in your field? Where did these darnel weeds come from?’

“He said to them, ‘An enemy has done this.’ “The servants asked him, ‘Do you want us to go and gather them up?’

“But he said, ‘No, lest perhaps while you gather up the darnel weeds, you root up the wheat with them.

Let both grow together until the harvest, and in the harvest time I will tell the reapers, “First, gather up the darnel weeds, and bind them in bundles to burn them; but gather the wheat into my barn.” ’ ”



\chapterornament
\section*{Page 16}

\lettrine{T}{hen Jesus sent the multitudes away, and went into the house. His disciples came to him, saying, “Explain to us the parable of the darnel weeds of the field.”}

He answered them, “He who sows the good seed is the Son of Man,

the field is the world; and the good seed, these are the children of the Kingdom; and the darnel weeds are the children of the evil one.

The enemy who sowed them is the devil. The harvest is the end of the age, and the reapers are angels.

As therefore the darnel weeds are gathered up and burned with fire; so will it be at the end of this age.

The Son of Man will send out his angels, and they will gather out of his Kingdom all things that cause stumbling, and those who do iniquity,

and will cast them into the furnace of fire. There will be the weeping and the gnashing of teeth.

Then the righteous will shine like the sun in the Kingdom of their Father. He who has ears to hear, let him hear.

“Again, the Kingdom of Heaven is like treasure hidden in the field, which a man found, and hid. In his joy, he goes and sells all that he has, and buys that field.

“Again, the Kingdom of Heaven is like a man who is a merchant seeking fine pearls,

who having found one pearl of great price, he went and sold all that he had, and bought it.

“Again, the Kingdom of Heaven is like a dragnet, that was cast into the sea, and gathered some fish of every kind,

which, when it was filled, they drew up on the beach. They sat down, and gathered the good into containers, but the bad they threw away.

So will it be in the end of the world. The angels will come and separate the wicked from among the righteous,

and will cast them into the furnace of fire. There will be the weeping and the gnashing of teeth.”

Jesus said to them, “Have you understood all these things?” They answered him, “Yes, Lord.”

He said to them, “Therefore every scribe who has been made a disciple in the Kingdom of Heaven is like a man who is a householder, who brings out of his treasure new and old things.”

He said, “God’s Kingdom is as if a man should cast seed on the earth,

and should sleep and rise night and day, and the seed should spring up and grow, though he doesn’t know how.

For the earth bears fruit: first the blade, then the ear, then the full grain in the ear.



\chapterornament
\section*{Page 17}

\lettrine{B}{ut when the fruit is ripe, immediately he puts in the sickle, because the harvest has come.”}

He said, “How will we liken God’s Kingdom? Or with what parable will we illustrate it?

It’s like a grain of mustard seed, which, when it is sown in the earth, though it is less than all the seeds that are on the earth,

yet when it is sown, grows up, and becomes greater than all the herbs, and puts out great branches, so that the birds of the sky can lodge under its shadow.”

With many such parables he spoke the word to them, as they were able to hear it.

Without a parable he didn’t speak to them; but privately to his own disciples he explained everything.

As they went on the way, a certain man said to him, “I want to follow you wherever you go, Lord.”

Jesus said to him, “The foxes have holes, and the birds of the sky have nests, but the Son of Man has no place to lay his head.”

He said to another, “Follow me!” But he said, “Lord, allow me first to go and bury my father.”

But Jesus said to him, “Leave the dead to bury their own dead, but you go and announce God’s Kingdom.”

Another also said, “I want to follow you, Lord, but first allow me to say good-bye to those who are at my house.”

But Jesus said to him, “No one, having put his hand to the plow, and looking back, is fit for God’s Kingdom.”

After these things he went out, and saw a tax collector named Levi sitting at the tax office, and said to him, “Follow me!”

He left everything, and rose up and followed him.

Levi made a great feast for him in his house. There was a great crowd of tax collectors and others who were reclining with them.

He was reclining at the table in his house, and many tax collectors and sinners sat down with Jesus and his disciples, for there were many, and they followed him.

The scribes and the Pharisees, when they saw that he was eating with the sinners and tax collectors, said to his disciples, “Why is it that he eats and drinks with tax collectors and sinners?”

When Jesus heard it, he said to them, “Those who are healthy have no need for a physician, but those who are sick. I came not to call the righteous, but sinners to repentance.”

He also told a parable to them. “No one puts a piece from a new garment on an old garment, or else he will tear the new, and also the piece from the new will not match the old.

No one puts new wine into old wine skins, or else the new wine will burst the skins, and it will be spilled, and the skins will be destroyed.



\chapterornament
\section*{Page 18}

\lettrine{B}{ut new wine must be put into fresh wine skins, and both are preserved.}

When Jesus had finished these parables, he departed from there.

Coming into his own country, he taught them in their synagogue, so that they were astonished, and said, “Where did this man get this wisdom, and these mighty works?

Isn’t this the carpenter’s son? Isn’t his mother called Mary, and his brothers, James, Joses, Simon, and Judas?

Aren’t all of his sisters with us? Where then did this man get all of these things?”

They were offended by him. But Jesus said to them, “A prophet is not without honor, except in his own country, and in his own house.”

But when he saw the multitudes, he was moved with compassion for them, because they were harassed and scattered, like sheep without a shepherd.

He called to himself the twelve, and began to send them out two by two; and he gave them authority over the unclean spirits.

Jesus sent these twelve out, and commanded them, saying, “Don’t go among the Gentiles, and don’t enter into any city of the Samaritans.

Rather, go to the lost sheep of the house of Israel.

Don’t take any gold, silver, or brass in your money belts.

Take no bag for your journey, neither two coats, nor sandals, nor staff: for the laborer is worthy of his food.

Into whatever city or village you enter, find out who in it is worthy; and stay there until you go on.

As you enter into the household, greet it.

If the household is worthy, let your peace come on it, but if it isn’t worthy, let your peace return to you.

Whoever doesn’t receive you, nor hear your words, as you go out of that house or that city, shake the dust off your feet.

Most certainly I tell you, it will be more tolerable for the land of Sodom and Gomorrah in the day of judgment than for that city.

“Behold, I send you out as sheep among wolves. Therefore be wise as serpents, and harmless as doves.

But beware of men: for they will deliver you up to councils, and in their synagogues they will scourge you.

Yes, and you will be brought before governors and kings for my sake, for a testimony to them and to the nations.



\chapterornament
\section*{Page 19}

\lettrine{B}{ut when they persecute you in this city, flee into the next, for most certainly I tell you, you will not have gone through the cities of Israel until the Son of Man has come.}

Therefore don’t be afraid of them, for there is nothing covered that will not be revealed; and hidden that will not be known.

What I tell you in the darkness, speak in the light; and what you hear whispered in the ear, proclaim on the housetops.

Don’t be afraid of those who kill the body, but are not able to kill the soul. Rather, fear him who is able to destroy both soul and body in Gehenna.

“Aren’t two sparrows sold for an assarion coin? Not one of them falls on the ground apart from your Father’s will,

but the very hairs of your head are all numbered.

Therefore don’t be afraid. You are of more value than many sparrows.

They went out and preached that people should repent.

The apostles gathered themselves together to Jesus, and they told him all things, whatever they had done, and whatever they had taught.

After these things, Jesus was walking in Galilee, for he wouldn’t walk in Judea, because the Jews sought to kill him.

They don’t eat when they come from the marketplace unless they bathe themselves, and there are many other things, which they have received to hold to: washings of cups, pitchers, bronze vessels, and couches.)

The Pharisees and the scribes asked him, “Why don’t your disciples walk according to the tradition of the elders, but eat their bread with unwashed hands?”

He called all the multitude to himself, and said to them, “Hear me, all of you, and understand.

There is nothing from outside of the man, that going into him can defile him; but the things which proceed out of the man are those that defile the man.

If anyone has ears to hear, let him hear!”

When he had entered into a house away from the multitude, his disciples asked him about the parable.

He said to them, “Are you also without understanding? Don’t you perceive that whatever goes into the man from outside can’t defile him,

because it doesn’t go into his heart, but into his stomach, then into the latrine, making all foods clean?”

He said, “That which proceeds out of the man, that defiles the man.

For from within, out of the hearts of men, proceed evil thoughts, adulteries, sexual sins, murders, thefts,



\chapterornament
\section*{Page 20}

\lettrine{c}{ovetings, wickedness, deceit, lustful desires, an evil eye, blasphemy, pride, and foolishness.}

All these evil things come from within, and defile the man.”

From there he arose, and went away into the borders of Tyre and Sidon. He entered into a house, and didn’t want anyone to know it, but he couldn’t escape notice.

In that hour the disciples came to Jesus, saying, “Who then is greatest in the Kingdom of Heaven?”

Jesus called a little child to himself, and set him in the middle of them,

and said, “Most certainly I tell you, unless you turn, and become as little children, you will in no way enter into the Kingdom of Heaven.

Whoever therefore humbles himself as this little child is the greatest in the Kingdom of Heaven.

“Woe to the world because of occasions of stumbling! For it must be that the occasions come, but woe to that person through whom the occasion comes!

If your hand or your foot causes you to stumble, cut it off, and cast it from you. It is better for you to enter into life maimed or crippled, rather than having two hands or two feet to be cast into the eternal fire.

If your eye causes you to stumble, pluck it out, and cast it from you. It is better for you to enter into life with one eye, rather than having two eyes to be cast into the Gehenna of fire.

“What do you think? If a man has one hundred sheep, and one of them goes astray, doesn’t he leave the ninety-nine, go to the mountains, and seek that which has gone astray?

If he finds it, most certainly I tell you, he rejoices over it more than over the ninety-nine which have not gone astray.

Even so it is not the will of your Father who is in heaven that one of these little ones should perish.

“If your brother sins against you, go, show him his fault between you and him alone. If he listens to you, you have gained back your brother.

But if he doesn’t listen, take one or two more with you, that at the mouth of two or three witnesses every word may be established.

If he refuses to listen to them, tell it to the assembly. If he refuses to hear the assembly also, let him be to you as a Gentile or a tax collector.

Then Peter came and said to him, “Lord, how often shall my brother sin against me, and I forgive him? Until seven times?”

Jesus said to him, “I don’t tell you until seven times, but, until seventy times seven.

Therefore the Kingdom of Heaven is like a certain king, who wanted to reconcile accounts with his servants.

When he had begun to reconcile, one was brought to him who owed him ten thousand talents.



\chapterornament
\section*{Page 21}

\lettrine{B}{ut because he couldn’t pay, his lord commanded him to be sold, with his wife, his children, and all that he had, and payment to be made.}

The servant therefore fell down and knelt before him, saying, ‘Lord, have patience with me, and I will repay you all!’

The lord of that servant, being moved with compassion, released him, and forgave him the debt.

“But that servant went out, and found one of his fellow servants, who owed him one hundred denarii, and he grabbed him, and took him by the throat, saying, ‘Pay me what you owe!’

“So his fellow servant fell down at his feet and begged him, saying, ‘Have patience with me, and I will repay you!’

He would not, but went and cast him into prison, until he should pay back that which was due.

So when his fellow servants saw what was done, they were exceedingly sorry, and came and told their lord all that was done.

Then his lord called him in, and said to him, ‘You wicked servant! I forgave you all that debt, because you begged me.

Shouldn’t you also have had mercy on your fellow servant, even as I had mercy on you?’

His lord was angry, and delivered him to the tormentors, until he should pay all that was due to him.

So my heavenly Father will also do to you, if you don’t each forgive your brother from your hearts for his misdeeds.”

Now after these things, the Lord also appointed seventy others, and sent them two by two ahead of him into every city and place where he was about to come.

Then he said to them, “The harvest is indeed plentiful, but the laborers are few. Pray therefore to the Lord of the harvest, that he may send out laborers into his harvest.

Go your ways. Behold, I send you out as lambs among wolves.

Carry no purse, nor wallet, nor sandals. Greet no one on the way.

Into whatever house you enter, first say, ‘Peace be to this house.’

If a son of peace is there, your peace will rest on him; but if not, it will return to you.

Remain in that same house, eating and drinking the things they give, for the laborer is worthy of his wages. Don’t go from house to house.

Into whatever city you enter, and they receive you, eat the things that are set before you.

But into whatever city you enter, and they don’t receive you, go out into its streets and say,



\chapterornament
\section*{Page 22}

\lettrine{‘}{Even the dust from your city that clings to us, we wipe off against you. Nevertheless know this, that God’s Kingdom has come near to you.’}

I tell you, it will be more tolerable in that day for Sodom than for that city.

Now the feast of the Jews, the Feast of Booths, was at hand.

His brothers therefore said to him, “Depart from here and go into Judea, that your disciples also may see your works which you do.

For no one does anything in secret while he seeks to be known openly. If you do these things, reveal yourself to the world.”

For even his brothers didn’t believe in him.

Jesus therefore said to them, “My time has not yet come, but your time is always ready.

The world can’t hate you, but it hates me, because I testify about it, that its works are evil.

You go up to the feast. I am not yet going up to this feast, because my time is not yet fulfilled.”

Having said these things to them, he stayed in Galilee.

But when his brothers had gone up to the feast, then he also went up, not publicly, but as it were in secret.

The Jews therefore sought him at the feast, and said, “Where is he?”

There was much murmuring among the multitudes concerning him. Some said, “He is a good man.” Others said, “Not so, but he leads the multitude astray.”

Yet no one spoke openly of him for fear of the Jews.

But when it was now the middle of the feast, Jesus went up into the temple and taught.

The Jews therefore marveled, saying, “How does this man know letters, having never been educated?”

Jesus therefore answered them, “My teaching is not mine, but his who sent me.

Didn’t Moses give you the law, and yet none of you keeps the law? Why do you seek to kill me?”

The multitude answered, “You have a demon! Who seeks to kill you?”

Jesus answered them, “I did one work and you all marvel because of it.



\chapterornament
\section*{Page 23}

\lettrine{M}{oses has given you circumcision (not that it is of Moses, but of the fathers), and on the Sabbath you circumcise a boy.}

If a boy receives circumcision on the Sabbath, that the law of Moses may not be broken, are you angry with me, because I made a man completely healthy on the Sabbath?

Don’t judge according to appearance, but judge righteous judgment.”

Therefore some of them of Jerusalem said, “Isn’t this he whom they seek to kill?

Behold, he speaks openly, and they say nothing to him. Can it be that the rulers indeed know that this is truly the Christ?

The Pharisees heard the multitude murmuring these things concerning him, and the chief priests and the Pharisees sent officers to arrest him.

So a division arose in the multitude because of him.

Some of them would have arrested him, but no one laid hands on him.

The officers therefore came to the chief priests and Pharisees, and they said to them, “Why didn’t you bring him?”

The officers answered, “No man ever spoke like this man!”

The Pharisees therefore answered them, “You aren’t also led astray, are you?

Have any of the rulers believed in him, or of the Pharisees?

But this multitude that doesn’t know the law is cursed.”

Nicodemus (he who came to him by night, being one of them) said to them,

“Does our law judge a man, unless it first hears from him personally and knows what he does?”

They answered him, “Are you also from Galilee? Search, and see that no prophet has arisen out of Galilee.”

Everyone went to his own house,

but Jesus went to the Mount of Olives.

Now very early in the morning, he came again into the temple, and all the people came to him. He sat down and taught them.

The scribes and the Pharisees brought a woman taken in adultery. Having set her in the middle,



\chapterornament
\section*{Page 24}

\lettrine{t}{hey told him, “Teacher, we found this woman in adultery, in the very act.}

Now in our law, Moses commanded us to stone such women. What then do you say about her?”

They said this testing him, that they might have something to accuse him of. But Jesus stooped down and wrote on the ground with his finger.

But when they continued asking him, he looked up and said to them, “He who is without sin among you, let him throw the first stone at her.”

Again he stooped down and wrote on the ground with his finger.

They, when they heard it, being convicted by their conscience, went out one by one, beginning from the oldest, even to the last. Jesus was left alone with the woman where she was, in the middle.

Jesus, standing up, saw her and said, “Woman, where are your accusers? Did no one condemn you?”

She said, “No one, Lord.” Jesus said, “Neither do I condemn you. Go your way. From now on, sin no more.”

As he passed by, he saw a man blind from birth.

His disciples asked him, “Rabbi, who sinned, this man or his parents, that he was born blind?”

Jesus answered, “This man didn’t sin, nor did his parents; but, that the works of God might be revealed in him.

“Most certainly, I tell you, one who doesn’t enter by the door into the sheep fold, but climbs up some other way, is a thief and a robber.

But one who enters in by the door is the shepherd of the sheep.

The gatekeeper opens the gate for him, and the sheep listen to his voice. He calls his own sheep by name, and leads them out.

Whenever he brings out his own sheep, he goes before them, and the sheep follow him, for they know his voice.

They will by no means follow a stranger, but will flee from him; for they don’t know the voice of strangers.”

I am the good shepherd. The good shepherd lays down his life for the sheep.

He who is a hired hand, and not a shepherd, who doesn’t own the sheep, sees the wolf coming, leaves the sheep, and flees. The wolf snatches the sheep, and scatters them.

The hired hand flees because he is a hired hand, and doesn’t care for the sheep.

I am the good shepherd. I know my own, and I’m known by my own;



\chapterornament
\section*{Page 25}

\lettrine{e}{ven as the Father knows me, and I know the Father. I lay down my life for the sheep.}

I have other sheep, which are not of this fold. I must bring them also, and they will hear my voice. They will become one flock with one shepherd.

Behold, a certain lawyer stood up and tested him, saying, “Teacher, what shall I do to inherit eternal life?”

He said to him, “What is written in the law? How do you read it?”

He answered, “You shall love the Lord your God with all your heart, with all your soul, with all your strength, and with all your mind; and your neighbor as yourself.”

He said to him, “You have answered correctly. Do this, and you will live.”

But he, desiring to justify himself, asked Jesus, “Who is my neighbor?”

Jesus answered, “A certain man was going down from Jerusalem to Jericho, and he fell among robbers, who both stripped him and beat him, and departed, leaving him half dead.

By chance a certain priest was going down that way. When he saw him, he passed by on the other side.

In the same way a Levite also, when he came to the place, and saw him, passed by on the other side.

But a certain Samaritan, as he traveled, came where he was. When he saw him, he was moved with compassion,

came to him, and bound up his wounds, pouring on oil and wine. He set him on his own animal, brought him to an inn, and took care of him.

On the next day, when he departed, he took out two denarii, gave them to the host, and said to him, ‘Take care of him. Whatever you spend beyond that, I will repay you when I return.’

Now which of these three do you think seemed to be a neighbor to him who fell among the robbers?”

He said, “He who showed mercy on him.” Then Jesus said to him, “Go and do likewise.”

When he finished praying in a certain place, one of his disciples said to him, “Lord, teach us to pray, just as John also taught his disciples.”

He said to them, “When you pray, say, ‘Our Father in heaven, may your name be kept holy. May your Kingdom come. May your will be done on earth, as it is in heaven.

Give us day by day our daily bread.

Forgive us our sins, for we ourselves also forgive everyone who is indebted to us. Bring us not into temptation, but deliver us from the evil one.’ ”

He said to them, “Which of you, if you go to a friend at midnight, and tell him, ‘Friend, lend me three loaves of bread,



\chapterornament
\section*{Page 26}

\lettrine{f}{or a friend of mine has come to me from a journey, and I have nothing to set before him,’}

and he from within will answer and say, ‘Don’t bother me. The door is now shut, and my children are with me in bed. I can’t get up and give it to you’?

I tell you, although he will not rise and give it to him because he is his friend, yet because of his persistence, he will get up and give him as many as he needs.

“I tell you, keep asking, and it will be given you. Keep seeking, and you will find. Keep knocking, and it will be opened to you.

For everyone who asks receives. He who seeks finds. To him who knocks it will be opened.

“Which of you fathers, if your son asks for bread, will give him a stone? Or if he asks for a fish, he won’t give him a snake instead of a fish, will he?

Or if he asks for an egg, he won’t give him a scorpion, will he?

If you then, being evil, know how to give good gifts to your children, how much more will your heavenly Father give the Holy Spirit to those who ask him?”

When he went into the house of one of the rulers of the Pharisees on a Sabbath to eat bread, they were watching him.

Behold, a certain man who had dropsy was in front of him.

Jesus, answering, spoke to the lawyers and Pharisees, saying, “Is it lawful to heal on the Sabbath?”

But they were silent. He took him, and healed him, and let him go.

He answered them, “Which of you, if your son or an ox fell into a well, wouldn’t immediately pull him out on a Sabbath day?”

They couldn’t answer him regarding these things.

He spoke a parable to those who were invited, when he noticed how they chose the best seats, and said to them,

“When you are invited by anyone to a wedding feast, don’t sit in the best seat, since perhaps someone more honorable than you might be invited by him,

and he who invited both of you would come and tell you, ‘Make room for this person.’ Then you would begin, with shame, to take the lowest place.

But when you are invited, go and sit in the lowest place, so that when he who invited you comes, he may tell you, ‘Friend, move up higher.’ Then you will be honored in the presence of all who sit at the table with you.

For everyone who exalts himself will be humbled, and whoever humbles himself will be exalted.”

He also said to the one who had invited him, “When you make a dinner or a supper, don’t call your friends, nor your brothers, nor your kinsmen, nor rich neighbors, or perhaps they might also return the favor, and pay you back.



\chapterornament
\section*{Page 27}

\lettrine{B}{ut when you make a feast, ask the poor, the maimed, the lame, or the blind;}

and you will be blessed, because they don’t have the resources to repay you. For you will be repaid in the resurrection of the righteous.”

But he said to him, “A certain man made a great supper, and he invited many people.

He sent out his servant at supper time to tell those who were invited, ‘Come, for everything is ready now.’

They all as one began to make excuses. “The first said to him, ‘I have bought a field, and I must go and see it. Please have me excused.’

“Another said, ‘I have bought five yoke of oxen, and I must go try them out. Please have me excused.’

“Another said, ‘I have married a wife, and therefore I can’t come.’

“That servant came, and told his lord these things. Then the master of the house, being angry, said to his servant, ‘Go out quickly into the streets and lanes of the city, and bring in the poor, maimed, blind, and lame.’

“The servant said, ‘Lord, it is done as you commanded, and there is still room.’

“The lord said to the servant, ‘Go out into the highways and hedges, and compel them to come in, that my house may be filled.

For I tell you that none of those men who were invited will taste of my supper.’ ”

For which of you, desiring to build a tower, doesn’t first sit down and count the cost, to see if he has enough to complete it?

Or perhaps, when he has laid a foundation, and is not able to finish, everyone who sees begins to mock him,

saying, ‘This man began to build, and wasn’t able to finish.’

Or what king, as he goes to encounter another king in war, will not sit down first and consider whether he is able with ten thousand to meet him who comes against him with twenty thousand?

Or else, while the other is yet a great way off, he sends an envoy, and asks for conditions of peace.

Now all the tax collectors and sinners were coming close to him to hear him.

The Pharisees and the scribes murmured, saying, “This man welcomes sinners, and eats with them.”

He told them this parable.

“Which of you men, if you had one hundred sheep, and lost one of them, wouldn’t leave the ninety-nine in the wilderness, and go after the one that was lost, until he found it?



\chapterornament
\section*{Page 28}

\lettrine{W}{hen he has found it, he carries it on his shoulders, rejoicing.}

When he comes home, he calls together his friends and his neighbors, saying to them, ‘Rejoice with me, for I have found my sheep which was lost!’

I tell you that even so there will be more joy in heaven over one sinner who repents, than over ninety-nine righteous people who need no repentance.

Or what woman, if she had ten drachma coins, if she lost one drachma coin, wouldn’t light a lamp, sweep the house, and seek diligently until she found it?

When she has found it, she calls together her friends and neighbors, saying, ‘Rejoice with me, for I have found the drachma which I had lost.’

Even so, I tell you, there is joy in the presence of the angels of God over one sinner repenting.”

He said, “A certain man had two sons.

The younger of them said to his father, ‘Father, give me my share of your property.’ He divided his livelihood between them.

Not many days after, the younger son gathered all of this together and traveled into a far country. There he wasted his property with riotous living.

When he had spent all of it, there arose a severe famine in that country, and he began to be in need.

He went and joined himself to one of the citizens of that country, and he sent him into his fields to feed pigs.

He wanted to fill his belly with the husks that the pigs ate, but no one gave him any.

But when he came to himself he said, ‘How many hired servants of my father’s have bread enough to spare, and I’m dying with hunger!

I will get up and go to my father, and will tell him, “Father, I have sinned against heaven, and in your sight.

I am no more worthy to be called your son. Make me as one of your hired servants.” ’

“He arose, and came to his father. But while he was still far off, his father saw him, and was moved with compassion, and ran, and fell on his neck, and kissed him.

The son said to him, ‘Father, I have sinned against heaven and in your sight. I am no longer worthy to be called your son.’

“But the father said to his servants, ‘Bring out the best robe, and put it on him. Put a ring on his hand, and sandals on his feet.

Bring the fattened calf, kill it, and let’s eat, and celebrate;

for this, my son, was dead, and is alive again. He was lost, and is found.’ Then they began to celebrate.



\chapterornament
\section*{Page 29}

\lettrine{“}{Now his elder son was in the field. As he came near to the house, he heard music and dancing.}

He called one of the servants to him, and asked what was going on.

He said to him, ‘Your brother has come, and your father has killed the fattened calf, because he has received him back safe and healthy.’

But he was angry, and would not go in. Therefore his father came out, and begged him.

But he answered his father, ‘Behold, these many years I have served you, and I never disobeyed a commandment of yours, but you never gave me a goat, that I might celebrate with my friends.

But when this your son came, who has devoured your living with prostitutes, you killed the fattened calf for him.’

“He said to him, ‘Son, you are always with me, and all that is mine is yours.

But it was appropriate to celebrate and be glad, for this, your brother, was dead, and is alive again. He was lost, and is found.’ ”

He also said to his disciples, “There was a certain rich man who had a manager. An accusation was made to him that this man was wasting his possessions.

He called him, and said to him, ‘What is this that I hear about you? Give an accounting of your management, for you can no longer be manager.’

“The manager said within himself, ‘What will I do, seeing that my lord is taking away the management position from me? I don’t have strength to dig. I am ashamed to beg.

I know what I will do, so that when I am removed from management, they may receive me into their houses.’

Calling each one of his lord’s debtors to him, he said to the first, ‘How much do you owe to my lord?’

He said, ‘A hundred batos of oil.’ He said to him, ‘Take your bill, and sit down quickly and write fifty.’

Then he said to another, ‘How much do you owe?’ He said, ‘A hundred cors of wheat.’ He said to him, ‘Take your bill, and write eighty.’

“His lord commended the dishonest manager because he had done wisely, for the children of this world are, in their own generation, wiser than the children of the light.

I tell you, make for yourselves friends by means of unrighteous mammon, so that when you fail, they may receive you into the eternal tents.

He who is faithful in a very little is faithful also in much. He who is dishonest in a very little is also dishonest in much.

If therefore you have not been faithful in the unrighteous mammon, who will commit to your trust the true riches?

If you have not been faithful in that which is another’s, who will give you that which is your own?



\chapterornament
\section*{Page 30}

\lettrine{N}{o servant can serve two masters, for either he will hate the one, and love the other; or else he will hold to one, and despise the other. You aren’t able to serve God and Mammon.”}

The Pharisees, who were lovers of money, also heard all these things, and they scoffed at him.

He said to them, “You are those who justify yourselves in the sight of men, but God knows your hearts. For that which is exalted among men is an abomination in the sight of God.

Everyone who divorces his wife and marries another commits adultery. He who marries one who is divorced from a husband commits adultery.

“Now there was a certain rich man, and he was clothed in purple and fine linen, living in luxury every day.

A certain beggar, named Lazarus, was taken to his gate, full of sores,

and desiring to be fed with the crumbs that fell from the rich man’s table. Yes, even the dogs came and licked his sores.

The beggar died, and he was carried away by the angels to Abraham’s bosom. The rich man also died, and was buried.

In Hades, he lifted up his eyes, being in torment, and saw Abraham far off, and Lazarus at his bosom.

He cried and said, ‘Father Abraham, have mercy on me, and send Lazarus, that he may dip the tip of his finger in water, and cool my tongue! For I am in anguish in this flame.’

“But Abraham said, ‘Son, remember that you, in your lifetime, received your good things, and Lazarus, in the same way, bad things. But here he is now comforted, and you are in anguish.

Besides all this, between us and you there is a great gulf fixed, that those who want to pass from here to you are not able, and that no one may cross over from there to us.’

“He said, ‘I ask you therefore, father, that you would send him to my father’s house;

for I have five brothers, that he may testify to them, so they won’t also come into this place of torment.’

“But Abraham said to him, ‘They have Moses and the prophets. Let them listen to them.’

“He said, ‘No, father Abraham, but if one goes to them from the dead, they will repent.’

“He said to him, ‘If they don’t listen to Moses and the prophets, neither will they be persuaded if one rises from the dead.’ ”

He said to the disciples, “It is impossible that no occasions of stumbling should come, but woe to him through whom they come!

It would be better for him if a millstone were hung around his neck, and he were thrown into the sea, rather than that he should cause one of these little ones to stumble.

Be careful. If your brother sins against you, rebuke him. If he repents, forgive him.



\chapterornament
\section*{Page 31}

\lettrine{I}{f he sins against you seven times in the day, and seven times returns, saying, ‘I repent,’ you shall forgive him.”}

But who is there among you, having a servant plowing or keeping sheep, that will say when he comes in from the field, ‘Come immediately and sit down at the table,’

and will not rather tell him, ‘Prepare my supper, clothe yourself properly, and serve me, while I eat and drink. Afterward you shall eat and drink’?

Does he thank that servant because he did the things that were commanded? I think not.

Even so you also, when you have done all the things that are commanded you, say, ‘We are unworthy servants. We have done our duty.’ ”

Being asked by the Pharisees when God’s Kingdom would come, he answered them, “God’s Kingdom doesn’t come with observation;

As it was in the days of Noah, even so will it be also in the days of the Son of Man.

They ate, they drank, they married, and they were given in marriage until the day that Noah entered into the ship, and the flood came and destroyed them all.

Likewise, even as it was in the days of Lot: they ate, they drank, they bought, they sold, they planted, they built;

but in the day that Lot went out from Sodom, it rained fire and sulfur from the sky and destroyed them all.

It will be the same way in the day that the Son of Man is revealed.

In that day, he who will be on the housetop and his goods in the house, let him not go down to take them away. Let him who is in the field likewise not turn back.

Remember Lot’s wife!

Whoever seeks to save his life loses it, but whoever loses his life preserves it.

I tell you, in that night there will be two people in one bed. One will be taken and the other will be left.

There will be two grinding grain together. One will be taken and the other will be left.”

He also spoke a parable to them that they must always pray, and not give up,

saying, “There was a judge in a certain city who didn’t fear God, and didn’t respect man.

A widow was in that city, and she often came to him, saying, ‘Defend me from my adversary!’

He wouldn’t for a while, but afterward he said to himself, ‘Though I neither fear God, nor respect man,



\chapterornament
\section*{Page 32}

\lettrine{y}{et because this widow bothers me, I will defend her, or else she will wear me out by her continual coming.’ ”}

The Lord said, “Listen to what the unrighteous judge says.

Won’t God avenge his chosen ones who are crying out to him day and night, and yet he exercises patience with them?

I tell you that he will avenge them quickly. Nevertheless, when the Son of Man comes, will he find faith on the earth?”

He spoke also this parable to certain people who were convinced of their own righteousness, and who despised all others.

“Two men went up into the temple to pray; one was a Pharisee, and the other was a tax collector.

The Pharisee stood and prayed to himself like this: ‘God, I thank you that I am not like the rest of men, extortionists, unrighteous, adulterers, or even like this tax collector.

I fast twice a week. I give tithes of all that I get.’

But the tax collector, standing far away, wouldn’t even lift up his eyes to heaven, but beat his breast, saying, ‘God, be merciful to me, a sinner!’

I tell you, this man went down to his house justified rather than the other; for everyone who exalts himself will be humbled, but he who humbles himself will be exalted.”

As they went on their way, he entered into a certain village, and a certain woman named Martha received him into her house.

She had a sister called Mary, who also sat at Jesus’ feet, and heard his word.

But Martha was distracted with much serving, and she came up to him, and said, “Lord, don’t you care that my sister left me to serve alone? Ask her therefore to help me.”

Jesus answered her, “Martha, Martha, you are anxious and troubled about many things,

but one thing is needed. Mary has chosen the good part, which will not be taken away from her.”

When Jesus had finished these words, he departed from Galilee, and came into the borders of Judea beyond the Jordan.

Great multitudes followed him, and he healed them there.

Pharisees came to him, testing him, and saying, “Is it lawful for a man to divorce his wife for any reason?”

He answered, “Haven’t you read that he who made them from the beginning made them male and female,

and said, ‘For this cause a man shall leave his father and mother, and shall be joined to his wife; and the two shall become one flesh?’



\chapterornament
\section*{Page 33}

\lettrine{S}{o that they are no more two, but one flesh. What therefore God has joined together, don’t let man tear apart.”}

They asked him, “Why then did Moses command us to give her a certificate of divorce, and divorce her?”

He said to them, “Moses, because of the hardness of your hearts, allowed you to divorce your wives, but from the beginning it has not been so.

I tell you that whoever divorces his wife, except for sexual immorality, and marries another, commits adultery; and he who marries her when she is divorced commits adultery.”

His disciples said to him, “If this is the case of the man with his wife, it is not expedient to marry.”

But he said to them, “Not all men can receive this saying, but those to whom it is given.

For there are eunuchs who were born that way from their mother’s womb, and there are eunuchs who were made eunuchs by men; and there are eunuchs who made themselves eunuchs for the Kingdom of Heaven’s sake. He who is able to receive it, let him receive it.”

Then little children were brought to him, that he should lay his hands on them and pray; and the disciples rebuked them.

But Jesus said, “Allow the little children, and don’t forbid them to come to me; for the Kingdom of Heaven belongs to ones like these.”

He laid his hands on them, and departed from there.

Behold, one came to him and said, “Good teacher, what good thing shall I do, that I may have eternal life?”

He said to him, “Why do you call me good? No one is good but one, that is, God. But if you want to enter into life, keep the commandments.”

He said to him, “Which ones?” Jesus said, “ ‘You shall not murder.’ ‘You shall not commit adultery.’ ‘You shall not steal.’ ‘You shall not offer false testimony.’

‘Honor your father and your mother.’ And, ‘You shall love your neighbor as yourself.’ ”

The young man said to him, “All these things I have observed from my youth. What do I still lack?”

Jesus said to him, “If you want to be perfect, go, sell what you have, and give to the poor, and you will have treasure in heaven; and come, follow me.”

But when the young man heard the saying, he went away sad, for he was one who had great possessions.

Jesus said to his disciples, “Most certainly I say to you, a rich man will enter into the Kingdom of Heaven with difficulty.

Again I tell you, it is easier for a camel to go through a needle’s eye, than for a rich man to enter into God’s Kingdom.”

When the disciples heard it, they were exceedingly astonished, saying, “Who then can be saved?”



\chapterornament
\section*{Page 34}

\lettrine{L}{ooking at them, Jesus said, “With men this is impossible, but with God all things are possible.”}

“For the Kingdom of Heaven is like a man who was the master of a household, who went out early in the morning to hire laborers for his vineyard.

When he had agreed with the laborers for a denarius a day, he sent them into his vineyard.

He went out about the third hour, and saw others standing idle in the marketplace.

He said to them, ‘You also go into the vineyard, and whatever is right I will give you.’ So they went their way.

Again he went out about the sixth and the ninth hour, and did likewise.

About the eleventh hour he went out, and found others standing idle. He said to them, ‘Why do you stand here all day idle?’

“They said to him, ‘Because no one has hired us.’ “He said to them, ‘You also go into the vineyard, and you will receive whatever is right.’

When evening had come, the lord of the vineyard said to his manager, ‘Call the laborers and pay them their wages, beginning from the last to the first.’

“When those who were hired at about the eleventh hour came, they each received a denarius.

When the first came, they supposed that they would receive more; and they likewise each received a denarius.

When they received it, they murmured against the master of the household,

saying, ‘These last have spent one hour, and you have made them equal to us, who have borne the burden of the day and the scorching heat!’

“But he answered one of them, ‘Friend, I am doing you no wrong. Didn’t you agree with me for a denarius?

Take that which is yours, and go your way. It is my desire to give to this last just as much as to you.

Isn’t it lawful for me to do what I want to with what I own? Or is your eye evil, because I am good?’

So the last will be first, and the first last. For many are called, but few are chosen.”

He entered and was passing through Jericho.

There was a man named Zacchaeus. He was a chief tax collector, and he was rich.

He was trying to see who Jesus was, and couldn’t because of the crowd, because he was short.



\chapterornament
\section*{Page 35}

\lettrine{H}{e ran on ahead, and climbed up into a sycamore tree to see him, for he was going to pass that way.}

When Jesus came to the place, he looked up and saw him, and said to him, “Zacchaeus, hurry and come down, for today I must stay at your house.”

He hurried, came down, and received him joyfully.

When they saw it, they all murmured, saying, “He has gone in to lodge with a man who is a sinner.”

Zacchaeus stood and said to the Lord, “Behold, Lord, half of my goods I give to the poor. If I have wrongfully exacted anything of anyone, I restore four times as much.”

Jesus said to him, “Today, salvation has come to this house, because he also is a son of Abraham.

For the Son of Man came to seek and to save that which was lost.”

As they heard these things, he went on and told a parable, because he was near Jerusalem, and they supposed that God’s Kingdom would be revealed immediately.

He said therefore, “A certain nobleman went into a far country to receive for himself a kingdom and to return.

He called ten servants of his and gave them ten mina coins,  and told them, ‘Conduct business until I come.’

But his citizens hated him, and sent an envoy after him, saying, ‘We don’t want this man to reign over us.’

“When he had come back again, having received the kingdom, he commanded these servants, to whom he had given the money, to be called to him, that he might know what they had gained by conducting business.

The first came before him, saying, ‘Lord, your mina has made ten more minas.’

“He said to him, ‘Well done, you good servant! Because you were found faithful with very little, you shall have authority over ten cities.’

“The second came, saying, ‘Your mina, Lord, has made five minas.’

“So he said to him, ‘And you are to be over five cities.’

Another came, saying, ‘Lord, behold, your mina, which I kept laid away in a handkerchief,

for I feared you, because you are an exacting man. You take up that which you didn’t lay down, and reap that which you didn’t sow.’

“He said to him, ‘Out of your own mouth I will judge you, you wicked servant! You knew that I am an exacting man, taking up that which I didn’t lay down, and reaping that which I didn’t sow.

Then why didn’t you deposit my money in the bank, and at my coming, I might have earned interest on it?’



\chapterornament
\section*{Page 36}

\lettrine{H}{e said to those who stood by, ‘Take the mina away from him and give it to him who has the ten minas.’}

“They said to him, ‘Lord, he has ten minas!’

‘For I tell you that to everyone who has, will more be given; but from him who doesn’t have, even that which he has will be taken away from him.

But bring those enemies of mine who didn’t want me to reign over them here, and kill them before me.’ ”

Having said these things, he went on ahead, going up to Jerusalem.

When they came near to Jerusalem, and came to Bethsphage, to the Mount of Olives, then Jesus sent two disciples,

saying to them, “Go into the village that is opposite you, and immediately you will find a donkey tied, and a colt with her. Untie them, and bring them to me.

If anyone says anything to you, you shall say, ‘The Lord needs them,’ and immediately he will send them.”

The disciples went, and did just as Jesus commanded them,

and brought the donkey and the colt, and laid their clothes on them; and he sat on them.

A very great multitude spread their clothes on the road. Others cut branches from the trees, and spread them on the road.

When he had come into Jerusalem, all the city was stirred up, saying, “Who is this?”

The Pharisees therefore said among themselves, “See how you accomplish nothing. Behold, the world has gone after him.”

Now there were certain Greeks among those that went up to worship at the feast.

These, therefore, came to Philip, who was from Bethsaida of Galilee, and asked him, saying, “Sir, we want to see Jesus.”

Philip came and told Andrew, and in turn, Andrew came with Philip, and they told Jesus.

Jesus answered them, “The time has come for the Son of Man to be glorified.

Most certainly I tell you, unless a grain of wheat falls into the earth and dies, it remains by itself alone. But if it dies, it bears much fruit.

He left them, and went out of the city to Bethany, and camped there.

The next day, when they had come out from Bethany, he was hungry.



\chapterornament
\section*{Page 37}

\lettrine{T}{hey came to Jerusalem, and Jesus entered into the temple, and began to throw out those who sold and those who bought in the temple, and overthrew the money changers’ tables, and the seats of those who sold the doves.}

He would not allow anyone to carry a container through the temple.

He taught, saying to them, “Isn’t it written, ‘My house will be called a house of prayer for all the nations?’ But you have made it a den of robbers!”

The chief priests and the scribes heard it, and sought how they might destroy him. For they feared him, because all the multitude was astonished at his teaching.

When evening came, he went out of the city.

They came again to Jerusalem, and as he was walking in the temple, the chief priests, the scribes, and the elders came to him,

But what do you think? A man had two sons, and he came to the first, and said, ‘Son, go work today in my vineyard.’

He answered, ‘I will not,’ but afterward he changed his mind, and went.

He came to the second, and said the same thing. He answered, ‘I’m going, sir,’ but he didn’t go.

Which of the two did the will of his father?” They said to him, “The first.” Jesus said to them, “Most certainly I tell you that the tax collectors and the prostitutes are entering into God’s Kingdom before you.

“Hear another parable. There was a man who was a master of a household, who planted a vineyard, set a hedge about it, dug a wine press in it, built a tower, leased it out to farmers, and went into another country.

He began to speak to them in parables. “A man planted a vineyard, put a hedge around it, dug a pit for the wine press, built a tower, rented it out to a farmer, and went into another country.

When it was time, he sent a servant to the farmer to get from the farmer his share of the fruit of the vineyard.

They took him, beat him, and sent him away empty.

Again, he sent another servant to them; and they threw stones at him, wounded him in the head, and sent him away shamefully treated.

Again he sent another; and they killed him; and many others, beating some, and killing some.

Therefore still having one, his beloved son, he sent him last to them, saying, ‘They will respect my son.’

But those farmers said among themselves, ‘This is the heir. Come, let’s kill him, and the inheritance will be ours.’

They took him, killed him, and cast him out of the vineyard.

What therefore will the lord of the vineyard do? He will come and destroy the farmers, and will give the vineyard to others.



\chapterornament
\section*{Page 38}

\lettrine{W}{hen the chief priests and the Pharisees heard his parables, they perceived that he spoke about them.}

When they sought to seize him, they feared the multitudes, because they considered him to be a prophet.

Jesus answered and spoke to them again in parables, saying,

“The Kingdom of Heaven is like a certain king, who made a wedding feast for his son,

and sent out his servants to call those who were invited to the wedding feast, but they would not come.

Again he sent out other servants, saying, ‘Tell those who are invited, “Behold, I have prepared my dinner. My cattle and my fatlings are killed, and all things are ready. Come to the wedding feast!” ’

But they made light of it, and went their ways, one to his own farm, another to his merchandise,

and the rest grabbed his servants, and treated them shamefully, and killed them.

When the king heard that, he was angry, and sent his armies, destroyed those murderers, and burned their city.

“Then he said to his servants, ‘The wedding is ready, but those who were invited weren’t worthy.

Go therefore to the intersections of the highways, and as many as you may find, invite to the wedding feast.’

Those servants went out into the highways, and gathered together as many as they found, both bad and good. The wedding was filled with guests.

But when the king came in to see the guests, he saw there a man who didn’t have on wedding clothing,

and he said to him, ‘Friend, how did you come in here not wearing wedding clothing?’ He was speechless.

Then the king said to the servants, ‘Bind him hand and foot, take him away, and throw him into the outer darkness. That is where the weeping and grinding of teeth will be.’

For many are called, but few chosen.”

Then the Pharisees went and took counsel how they might entrap him in his talk.

They sent their disciples to him, along with the Herodians, saying, “Teacher, we know that you are honest, and teach the way of God in truth, no matter whom you teach, for you aren’t partial to anyone.

Tell us therefore, what do you think? Is it lawful to pay taxes to Caesar, or not?”

But Jesus perceived their wickedness, and said, “Why do you test me, you hypocrites?



\chapterornament
\section*{Page 39}

\lettrine{S}{how me the tax money.” They brought to him a denarius.}

He asked them, “Whose is this image and inscription?”

They said to him, “Caesar’s.” Then he said to them, “Give therefore to Caesar the things that are Caesar’s, and to God the things that are God’s.”

When they heard it, they marveled, and left him, and went away.

On that day Sadducees (those who say that there is no resurrection) came to him. They asked him,

saying, “Teacher, Moses said, ‘If a man dies, having no children, his brother shall marry his wife, and raise up offspring for his brother.’

Now there were with us seven brothers. The first married and died, and having no offspring left his wife to his brother.

In the same way, the second also, and the third, to the seventh.

After them all, the woman died.

In the resurrection therefore, whose wife will she be of the seven? For they all had her.”

But Jesus answered them, “You are mistaken, not knowing the Scriptures, nor the power of God.

For in the resurrection they neither marry, nor are given in marriage, but are like God’s angels in heaven.

But concerning the resurrection of the dead, haven’t you read that which was spoken to you by God, saying,

‘I am the God of Abraham, and the God of Isaac, and the God of Jacob?’ God is not the God of the dead, but of the living.”

When the multitudes heard it, they were astonished at his teaching.

One of the scribes came, and heard them questioning together, and knowing that he had answered them well, asked him, “Which commandment is the greatest of all?”

Jesus answered, “The greatest is, ‘Hear, Israel, the Lord our God, the Lord is one:

you shall love the Lord your God with all your heart, and with all your soul, and with all your mind, and with all your strength.’ This is the first commandment.

The second is like this, ‘You shall love your neighbor as yourself.’ There is no other commandment greater than these.”

The whole law and the prophets depend on these two commandments.”



\chapterornament
\section*{Page 40}

\lettrine{T}{he scribe said to him, “Truly, teacher, you have said well that he is one, and there is none other but he,}

and to love him with all the heart, and with all the understanding, with all the soul, and with all the strength, and to love his neighbor as himself, is more important than all whole burnt offerings and sacrifices.”

Then Jesus spoke to the multitudes and to his disciples,

saying, “The scribes and the Pharisees sat on Moses’ seat.

All things therefore whatever they tell you to observe, observe and do, but don’t do their works; for they say, and don’t do.

For they bind heavy burdens that are grievous to be borne, and lay them on men’s shoulders; but they themselves will not lift a finger to help them.

But they do all their works to be seen by men. They make their phylacteries  broad, enlarge the fringes of their garments,

and love the place of honor at feasts, the best seats in the synagogues,

the salutations in the marketplaces, and to be called ‘Rabbi, Rabbi’ by men.

But don’t you be called ‘Rabbi,’ for one is your teacher, the Christ, and all of you are brothers.

Call no man on the earth your father, for one is your Father, he who is in heaven.

Neither be called masters, for one is your master, the Christ.

But he who is greatest among you will be your servant.

Whoever exalts himself will be humbled, and whoever humbles himself will be exalted.

“Woe to you, scribes and Pharisees, hypocrites! For you devour widows’ houses, and as a pretense you make long prayers. Therefore you will receive greater condemnation.

“But woe to you, scribes and Pharisees, hypocrites! Because you shut up the Kingdom of Heaven against men; for you don’t enter in yourselves, neither do you allow those who are entering in to enter.

Woe to you, scribes and Pharisees, hypocrites! For you travel around by sea and land to make one proselyte; and when he becomes one, you make him twice as much a son of Gehenna as yourselves.

“Woe to you, you blind guides, who say, ‘Whoever swears by the temple, it is nothing; but whoever swears by the gold of the temple, he is obligated.’

You blind fools! For which is greater, the gold, or the temple that sanctifies the gold?

‘Whoever swears by the altar, it is nothing; but whoever swears by the gift that is on it, he is obligated?’



\chapterornament
\section*{Page 41}

\lettrine{Y}{ou blind fools! For which is greater, the gift, or the altar that sanctifies the gift?}

He therefore who swears by the altar, swears by it, and by everything on it.

He who swears by the temple, swears by it, and by him who has been living in it.

He who swears by heaven, swears by the throne of God, and by him who sits on it.

“Woe to you, scribes and Pharisees, hypocrites! For you tithe mint, dill, and cumin, and have left undone the weightier matters of the law: justice, mercy, and faith. But you ought to have done these, and not to have left the other undone.

You blind guides, who strain out a gnat, and swallow a camel!

“Woe to you, scribes and Pharisees, hypocrites! For you clean the outside of the cup and of the platter, but within they are full of extortion and unrighteousness.

You blind Pharisee, first clean the inside of the cup and of the platter, that its outside may become clean also.

“Woe to you, scribes and Pharisees, hypocrites! For you are like whitened tombs, which outwardly appear beautiful, but inwardly are full of dead men’s bones, and of all uncleanness.

Even so you also outwardly appear righteous to men, but inwardly you are full of hypocrisy and iniquity.

“Woe to you, scribes and Pharisees, hypocrites! For you build the tombs of the prophets, and decorate the tombs of the righteous,

and say, ‘If we had lived in the days of our fathers, we wouldn’t have been partakers with them in the blood of the prophets.’

Therefore you testify to yourselves that you are children of those who killed the prophets.

Fill up, then, the measure of your fathers.

You serpents, you offspring of vipers, how will you escape the judgment of Gehenna?

Jesus sat down opposite the treasury, and saw how the multitude cast money into the treasury. Many who were rich cast in much.

A poor widow came, and she cast in two small brass coins, which equal a quadrans coin.

He called his disciples to himself, and said to them, “Most certainly I tell you, this poor widow gave more than all those who are giving into the treasury,

for they all gave out of their abundance, but she, out of her poverty, gave all that she had to live on.”

Jesus went out from the temple, and was going on his way. His disciples came to him to show him the buildings of the temple.



\chapterornament
\section*{Page 42}

\lettrine{B}{ut he answered them, “You see all of these things, don’t you? Most certainly I tell you, there will not be left here one stone on another, that will not be thrown down.”}

then let those who are in Judea flee to the mountains.

Let him who is on the housetop not go down to take out the things that are in his house.

Let him who is in the field not return back to get his clothes.

But woe to those who are with child and to nursing mothers in those days!

Pray that your flight will not be in the winter, nor on a Sabbath,

for then there will be great suffering, such as has not been from the beginning of the world until now, no, nor ever will be.

But immediately after the suffering of those days, the sun will be darkened, the moon will not give its light, the stars will fall from the sky, and the powers of the heavens will be shaken;

“Now from the fig tree learn this parable. When its branch has now become tender, and produces its leaves, you know that the summer is near.

Even so you also, when you see all these things, know that it is near, even at the doors.

But no one knows of that day and hour, not even the angels of heaven, but my Father only.

“As the days of Noah were, so will the coming of the Son of Man be.

For as in those days which were before the flood they were eating and drinking, marrying and giving in marriage, until the day that Noah entered into the ship,

and they didn’t know until the flood came, and took them all away, so will the coming of the Son of Man be.

Then two men will be in the field: one will be taken and one will be left.

Two women will be grinding at the mill: one will be taken and one will be left.

Watch therefore, for you don’t know in what hour your Lord comes.

But know this, that if the master of the house had known in what watch of the night the thief was coming, he would have watched, and would not have allowed his house to be broken into.

Therefore also be ready, for in an hour that you don’t expect, the Son of Man will come.

“Who then is the faithful and wise servant, whom his lord has set over his household, to give them their food in due season?



\chapterornament
\section*{Page 43}

\lettrine{B}{lessed is that servant whom his lord finds doing so when he comes.}

Most certainly I tell you that he will set him over all that he has.

But if that evil servant should say in his heart, ‘My lord is delaying his coming,’

and begins to beat his fellow servants, and eat and drink with the drunkards,

the lord of that servant will come in a day when he doesn’t expect it, and in an hour when he doesn’t know it,

and will cut him in pieces, and appoint his portion with the hypocrites. That is where the weeping and grinding of teeth will be.

“Then the Kingdom of Heaven will be like ten virgins, who took their lamps, and went out to meet the bridegroom.

Five of them were foolish, and five were wise.

Those who were foolish, when they took their lamps, took no oil with them,

but the wise took oil in their vessels with their lamps.

Now while the bridegroom delayed, they all slumbered and slept.

But at midnight there was a cry, ‘Behold! The bridegroom is coming! Come out to meet him!’

Then all those virgins arose, and trimmed their lamps.

The foolish said to the wise, ‘Give us some of your oil, for our lamps are going out.’

But the wise answered, saying, ‘What if there isn’t enough for us and you? You go rather to those who sell, and buy for yourselves.’

While they went away to buy, the bridegroom came, and those who were ready went in with him to the wedding feast, and the door was shut.

Afterward the other virgins also came, saying, ‘Lord, Lord, open to us.’

But he answered, ‘Most certainly I tell you, I don’t know you.’

Watch therefore, for you don’t know the day nor the hour in which the Son of Man is coming.

“For it is like a man, going into another country, who called his own servants, and entrusted his goods to them.



\chapterornament
\section*{Page 44}

\lettrine{T}{o one he gave five talents, to another two, to another one; to each according to his own ability. Then he went on his journey.}

Immediately he who received the five talents went and traded with them, and made another five talents.

In the same way, he also who got the two gained another two.

But he who received the one talent went away and dug in the earth, and hid his lord’s money.

“Now after a long time the lord of those servants came, and reconciled accounts with them.

He who received the five talents came and brought another five talents, saying, ‘Lord, you delivered to me five talents. Behold, I have gained another five talents in addition to them.’

“His lord said to him, ‘Well done, good and faithful servant. You have been faithful over a few things, I will set you over many things. Enter into the joy of your lord.’

“He also who got the two talents came and said, ‘Lord, you delivered to me two talents. Behold, I have gained another two talents in addition to them.’

“His lord said to him, ‘Well done, good and faithful servant. You have been faithful over a few things, I will set you over many things. Enter into the joy of your lord.’

“He also who had received the one talent came and said, ‘Lord, I knew you that you are a hard man, reaping where you didn’t sow, and gathering where you didn’t scatter.

I was afraid, and went away and hid your talent in the earth. Behold, you have what is yours.’

“But his lord answered him, ‘You wicked and slothful servant. You knew that I reap where I didn’t sow, and gather where I didn’t scatter.

You ought therefore to have deposited my money with the bankers, and at my coming I should have received back my own with interest.

Take away therefore the talent from him, and give it to him who has the ten talents.

For to everyone who has will be given, and he will have abundance, but from him who doesn’t have, even that which he has will be taken away.

Throw out the unprofitable servant into the outer darkness, where there will be weeping and gnashing of teeth.’

“So be careful, or your hearts will be loaded down with carousing, drunkenness, and cares of this life, and that day will come on you suddenly.

For it will come like a snare on all those who dwell on the surface of all the earth.

Therefore be watchful all the time, praying that you may be counted worthy to escape all these things that will happen, and to stand before the Son of Man.”

“But when the Son of Man comes in his glory, and all the holy angels with him, then he will sit on the throne of his glory.



\chapterornament
\section*{Page 45}

\lettrine{B}{efore him all the nations will be gathered, and he will separate them one from another, as a shepherd separates the sheep from the goats.}

He will set the sheep on his right hand, but the goats on the left.

Then the King will tell those on his right hand, ‘Come, blessed of my Father, inherit the Kingdom prepared for you from the foundation of the world;

for I was hungry, and you gave me food to eat. I was thirsty, and you gave me drink. I was a stranger, and you took me in.

I was naked, and you clothed me. I was sick, and you visited me. I was in prison, and you came to me.’

“Then the righteous will answer him, saying, ‘Lord, when did we see you hungry, and feed you; or thirsty, and give you a drink?

When did we see you as a stranger, and take you in; or naked, and clothe you?

When did we see you sick, or in prison, and come to you?’

“The King will answer them, ‘Most certainly I tell you, because you did it to one of the least of these my brothers,  you did it to me.’

Then he will say also to those on the left hand, ‘Depart from me, you cursed, into the eternal fire which is prepared for the devil and his angels;

for I was hungry, and you didn’t give me food to eat; I was thirsty, and you gave me no drink;

I was a stranger, and you didn’t take me in; naked, and you didn’t clothe me; sick, and in prison, and you didn’t visit me.’

“Then they will also answer, saying, ‘Lord, when did we see you hungry, or thirsty, or a stranger, or naked, or sick, or in prison, and didn’t help you?’

“Then he will answer them, saying, ‘Most certainly I tell you, because you didn’t do it to one of the least of these, you didn’t do it to me.’

These will go away into eternal punishment, but the righteous into eternal life.”

It was now two days before the feast of the Passover and the unleavened bread, and the chief priests and the scribes sought how they might seize him by deception, and kill him.

For they said, “Not during the feast, because there might be a riot among the people.”

While he was at Bethany, in the house of Simon the leper, as he sat at the table, a woman came having an alabaster jar of ointment of pure nard—very costly. She broke the jar, and poured it over his head.

But there were some who were indignant among themselves, saying, “Why has this ointment been wasted?

For this might have been sold for more than three hundred denarii,  and given to the poor.” So they grumbled against her.



\chapterornament
\section*{Page 46}

\lettrine{B}{ut Jesus said, “Leave her alone. Why do you trouble her? She has done a good work for me.}

For you always have the poor with you, and whenever you want to, you can do them good; but you will not always have me.

She has done what she could. She has anointed my body beforehand for the burying.

Then one of the twelve, who was called Judas Iscariot, went to the chief priests,

and said, “What are you willing to give me, that I should deliver him to you?” They weighed out for him thirty pieces of silver.

From that time he sought opportunity to betray him.

Now on the first day of unleavened bread, the disciples came to Jesus, saying to him, “Where do you want us to prepare for you to eat the Passover?”

He said, “Go into the city to a certain person, and tell him, ‘The Teacher says, “My time is at hand. I will keep the Passover at your house with my disciples.” ’ ”

The disciples did as Jesus commanded them, and they prepared the Passover.

Now when evening had come, he was reclining at the table with the twelve disciples.

A dispute also arose among them, which of them was considered to be greatest.

He said to them, “The kings of the nations lord it over them, and those who have authority over them are called ‘benefactors.’

But not so with you. But one who is the greater among you, let him become as the younger, and one who is governing, as one who serves.

For who is greater, one who sits at the table, or one who serves? Isn’t it he who sits at the table? But I am among you as one who serves.

During supper, the devil having already put into the heart of Judas Iscariot, Simon’s son, to betray him,

arose from supper, and laid aside his outer garments. He took a towel and wrapped a towel around his waist.

Then he poured water into the basin, and began to wash the disciples’ feet and to wipe them with the towel that was wrapped around him.

Then he came to Simon Peter. He said to him, “Lord, do you wash my feet?”

Jesus answered him, “You don’t know what I am doing now, but you will understand later.”

Peter said to him, “You will never wash my feet!” Jesus answered him, “If I don’t wash you, you have no part with me.”



\chapterornament
\section*{Page 47}

\lettrine{S}{imon Peter said to him, “Lord, not my feet only, but also my hands and my head!”}

Jesus said to him, “Someone who has bathed only needs to have his feet washed, but is completely clean. You are clean, but not all of you.”

For he knew him who would betray him, therefore he said, “You are not all clean.”

So when he had washed their feet, put his outer garment back on, and sat down again, he said to them, “Do you know what I have done to you?

You call me, ‘Teacher’ and ‘Lord.’ You say so correctly, for so I am.

If I then, the Lord and the Teacher, have washed your feet, you also ought to wash one another’s feet.

For I have given you an example, that you should also do as I have done to you.

Most certainly I tell you, a servant is not greater than his lord, neither is one who is sent greater than he who sent him.

If you know these things, blessed are you if you do them.

When Jesus had said this, he was troubled in spirit, and testified, “Most certainly I tell you that one of you will betray me.”

The disciples looked at one another, perplexed about whom he spoke.

One of his disciples, whom Jesus loved, was at the table, leaning against Jesus’ breast.

Simon Peter therefore beckoned to him, and said to him, “Tell us who it is of whom he speaks.”

He, leaning back, as he was, on Jesus’ breast, asked him, “Lord, who is it?”

Jesus therefore answered, “It is he to whom I will give this piece of bread when I have dipped it.” So when he had dipped the piece of bread, he gave it to Judas, the son of Simon Iscariot.

When he had gone out, Jesus said, “Now the Son of Man has been glorified, and God has been glorified in him.

A new commandment I give to you, that you love one another. Just as I have loved you, you also love one another.

By this everyone will know that you are my disciples, if you have love for one another.”

Then Jesus said to them, “All of you will be made to stumble because of me tonight, for it is written, ‘I will strike the shepherd, and the sheep of the flock will be scattered.’

But Peter answered him, “Even if all will be made to stumble because of you, I will never be made to stumble.”



\chapterornament
\section*{Page 48}

\lettrine{H}{e said to him, “Lord, I am ready to go with you both to prison and to death!”}

He said, “I tell you, Peter, the rooster will by no means crow today until you deny that you know me three times.”

Peter said to him, “Even if I must die with you, I will not deny you.” All of the disciples also said likewise.

Then Jesus came with them to a place called Gethsemane, and said to his disciples, “Sit here, while I go there and pray.”

He took with him Peter and the two sons of Zebedee, and began to be sorrowful and severely troubled.

Then he said to them, “My soul is exceedingly sorrowful, even to death. Stay here, and watch with me.”

He went forward a little, fell on his face, and prayed, saying, “My Father, if it is possible, let this cup pass away from me; nevertheless, not what I desire, but what you desire.”

He came to the disciples, and found them sleeping, and said to Peter, “What, couldn’t you watch with me for one hour?

Watch and pray, that you don’t enter into temptation. The spirit indeed is willing, but the flesh is weak.”

Again, a second time he went away, and prayed, saying, “My Father, if this cup can’t pass away from me unless I drink it, your desire be done.”

He came again and found them sleeping, for their eyes were heavy.

He left them again, went away, and prayed a third time, saying the same words.

Then he came to his disciples, and said to them, “Sleep on now, and take your rest. Behold, the hour is at hand, and the Son of Man is betrayed into the hands of sinners.

When Jesus had spoken these words, he went out with his disciples over the brook Kidron, where there was a garden, into which he and his disciples entered.

Now Judas, who betrayed him, also knew the place, for Jesus often met there with his disciples.

Judas then, having taken a detachment of soldiers and officers from the chief priests and the Pharisees, came there with lanterns, torches, and weapons.

Now he who betrayed him gave them a sign, saying, “Whoever I kiss, he is the one. Seize him.”

Immediately he came to Jesus, and said, “Hail, Rabbi!” and kissed him.

Jesus said to him, “Friend, why are you here?” Then they came and laid hands on Jesus, and took him.

Jesus therefore, knowing all the things that were happening to him, went out, and said to them, “Who are you looking for?”



\chapterornament
\section*{Page 49}

\lettrine{T}{hey answered him, “Jesus of Nazareth.” Jesus said to them, “I am he.” Judas also, who betrayed him, was standing with them.}

When therefore he said to them, “I am he,” they went backward, and fell to the ground.

Again therefore he asked them, “Who are you looking for?” They said, “Jesus of Nazareth.”

Jesus answered, “I told you that I am he. If therefore you seek me, let these go their way,”

Jesus said to him, “Friend, why are you here?” Then they came and laid hands on Jesus, and took him.

Behold, one of those who were with Jesus stretched out his hand, and drew his sword, and struck the servant of the high priest, and struck off his ear.

Then Jesus said to him, “Put your sword back into its place, for all those who take the sword will die by the sword.

In that hour Jesus said to the multitudes, “Have you come out as against a robber with swords and clubs to seize me? I sat daily in the temple teaching, and you didn’t arrest me.

But all this has happened that the Scriptures of the prophets might be fulfilled.” Then all the disciples left him and fled.

A certain young man followed him, having a linen cloth thrown around himself over his naked body. The young men grabbed him,

but he left the linen cloth, and fled from them naked.

Those who had taken Jesus led him away to Caiaphas the high priest, where the scribes and the elders were gathered together.

Simon Peter followed Jesus, as did another disciple. Now that disciple was known to the high priest, and entered in with Jesus into the court of the high priest;

but Peter was standing at the door outside. So the other disciple, who was known to the high priest, went out and spoke to her who kept the door, and brought in Peter.

Now the servants and the officers were standing there, having made a fire of coals, for it was cold. They were warming themselves. Peter was with them, standing and warming himself.

Then the maid who kept the door said to Peter, “Are you also one of this man’s disciples?” He said, “I am not.”

Now Simon Peter was standing and warming himself. They said therefore to him, “You aren’t also one of his disciples, are you?” He denied it and said, “I am not.”

One of the servants of the high priest, being a relative of him whose ear Peter had cut off, said, “Didn’t I see you in the garden with him?”

Peter therefore denied it again, and immediately the rooster crowed.

Peter remembered the word which Jesus had said to him, “Before the rooster crows, you will deny me three times.” Then he went out and wept bitterly.



\chapterornament
\section*{Page 50}

\lettrine{T}{he high priest therefore asked Jesus about his disciples and about his teaching.}

Jesus answered him, “I spoke openly to the world. I always taught in synagogues, and in the temple, where the Jews always meet. I said nothing in secret.

Why do you ask me? Ask those who have heard me what I said to them. Behold, they know the things which I said.”

When he had said this, one of the officers standing by slapped Jesus with his hand, saying, “Do you answer the high priest like that?”

Jesus answered him, “If I have spoken evil, testify of the evil; but if well, why do you beat me?”

They led Jesus away to the high priest. All the chief priests, the elders, and the scribes came together with him.

Now the chief priests and the whole council sought witnesses against Jesus to put him to death, and found none.

For many gave false testimony against him, and their testimony didn’t agree with each other.

Some stood up, and gave false testimony against him, saying,

“We heard him say, ‘I will destroy this temple that is made with hands, and in three days I will build another made without hands.’ ”

Even so, their testimony didn’t agree.

The high priest stood up in the middle, and asked Jesus, “Have you no answer? What is it which these testify against you?”

But he stayed quiet, and answered nothing. Again the high priest asked him, “Are you the Christ, the Son of the Blessed?”

“If you are the Christ, tell us.” But he said to them, “If I tell you, you won’t believe,

and if I ask, you will in no way answer me or let me go.

They all said, “Are you then the Son of God?” He said to them, “You say it, because I am.”

The high priest tore his clothes, and said, “What further need have we of witnesses?

You have heard the blasphemy! What do you think?” They all condemned him to be worthy of death.

Some began to spit on him, and to cover his face, and to beat him with fists, and to tell him, “Prophesy!” The officers struck him with the palms of their hands.

They led Jesus therefore from Caiaphas into the Praetorium. It was early, and they themselves didn’t enter into the Praetorium, that they might not be defiled, but might eat the Passover.



\chapterornament
\section*{Page 51}

\lettrine{P}{ilate therefore went out to them, and said, “What accusation do you bring against this man?”}

They answered him, “If this man weren’t an evildoer, we wouldn’t have delivered him up to you.”

Pilate therefore said to them, “Take him yourselves, and judge him according to your law.” Therefore the Jews said to him, “It is illegal for us to put anyone to death,”

Pilate therefore entered again into the Praetorium, called Jesus, and said to him, “Are you the King of the Jews?”

Jesus answered him, “Do you say this by yourself, or did others tell you about me?”

Pilate answered, “I’m not a Jew, am I? Your own nation and the chief priests delivered you to me. What have you done?”

Jesus answered, “My Kingdom is not of this world. If my Kingdom were of this world, then my servants would fight, that I wouldn’t be delivered to the Jews. But now my Kingdom is not from here.”

Pilate therefore said to him, “Are you a king then?” Jesus answered, “You say that I am a king. For this reason I have been born, and for this reason I have come into the world, that I should testify to the truth. Everyone who is of the truth listens to my voice.”

Pilate said to him, “What is truth?” When he had said this, he went out again to the Jews, and said to them, “I find no basis for a charge against him.

But they insisted, saying, “He stirs up the people, teaching throughout all Judea, beginning from Galilee even to this place.”

Then Pilate said to him, “Don’t you hear how many things they testify against you?”

But when Pilate heard Galilee mentioned, he asked if the man was a Galilean.

When he found out that he was in Herod’s jurisdiction, he sent him to Herod, who was also in Jerusalem during those days.

Now when Herod saw Jesus, he was exceedingly glad, for he had wanted to see him for a long time, because he had heard many things about him. He hoped to see some miracle done by him.

He questioned him with many words, but he gave no answers.

The chief priests and the scribes stood, vehemently accusing him.

Herod with his soldiers humiliated him and mocked him. Dressing him in luxurious clothing, they sent him back to Pilate.

Herod and Pilate became friends with each other that very day, for before that they were enemies with each other.

Pilate called together the chief priests, the rulers, and the people,

and said to them, “You brought this man to me as one that perverts the people, and behold, having examined him before you, I found no basis for a charge against this man concerning those things of which you accuse him.



\chapterornament
\section*{Page 52}

\lettrine{N}{either has Herod, for I sent you to him, and see, nothing worthy of death has been done by him.}

I will therefore chastise him and release him.”

Now at the feast the governor was accustomed to release to the multitude one prisoner, whom they desired.

They had then a notable prisoner, called Barabbas.

When therefore they were gathered together, Pilate said to them, “Whom do you want me to release to you? Barabbas, or Jesus, who is called Christ?”

For he knew that because of envy they had delivered him up.

While he was sitting on the judgment seat, his wife sent to him, saying, “Have nothing to do with that righteous man, for I have suffered many things today in a dream because of him.”

Now the chief priests and the elders persuaded the multitudes to ask for Barabbas, and destroy Jesus.

But the governor answered them, “Which of the two do you want me to release to you?” They said, “Barabbas!”

Pilate said to them, “What then shall I do to Jesus, who is called Christ?” They all said to him, “Let him be crucified!”

But the governor said, “Why? What evil has he done?” But they cried out exceedingly, saying, “Let him be crucified!”

Then he released to them Barabbas, but Jesus he flogged and delivered to be crucified.

Then the governor’s soldiers took Jesus into the Praetorium, and gathered the whole garrison together against him.

They braided a crown of thorns and put it on his head, and a reed in his right hand; and they kneeled down before him, and mocked him, saying, “Hail, King of the Jews!”

They spat on him, and took the reed and struck him on the head.

When they had mocked him, they took the robe off him, and put his clothes on him, and led him away to crucify him.

Then Judas, who betrayed him, when he saw that Jesus was condemned, felt remorse, and brought back the thirty pieces of silver to the chief priests and elders,

saying, “I have sinned in that I betrayed innocent blood.” But they said, “What is that to us? You see to it.”

He threw down the pieces of silver in the sanctuary, and departed. He went away and hanged himself.

The chief priests took the pieces of silver, and said, “It’s not lawful to put them into the treasury, since it is the price of blood.”



\chapterornament
\section*{Page 53}

\lettrine{T}{hey took counsel, and bought the potter’s field with them, to bury strangers in.}

Therefore that field was called “The Field of Blood” to this day.

When they led him away, they grabbed one Simon of Cyrene, coming from the country, and laid on him the cross, to carry it after Jesus.

A great multitude of the people followed him, including women who also mourned and lamented him.

But Jesus, turning to them, said, “Daughters of Jerusalem, don’t weep for me, but weep for yourselves and for your children.

For behold, the days are coming in which they will say, ‘Blessed are the barren, the wombs that never bore, and the breasts that never nursed.’

Then they will begin to tell the mountains, ‘Fall on us!’ and tell the hills, ‘Cover us.’

For if they do these things in the green tree, what will be done in the dry?”

There were also others, two criminals, led with him to be put to death.

He went out, bearing his cross, to the place called “The Place of a Skull”, which is called in Hebrew, “Golgotha”,

where they crucified him, and with him two others, on either side one, and Jesus in the middle.

Pilate wrote a title also, and put it on the cross. There was written, “JESUS OF NAZARETH, THE KING OF THE JEWS.”

Therefore many of the Jews read this title, for the place where Jesus was crucified was near the city; and it was written in Hebrew, in Latin, and in Greek.

The chief priests of the Jews therefore said to Pilate, “Don’t write, ‘The King of the Jews,’ but, ‘he said, “I am King of the Jews.” ’ ”

Pilate answered, “What I have written, I have written.”

Then the soldiers, when they had crucified Jesus, took his garments and made four parts, to every soldier a part; and also the coat. Now the coat was without seam, woven from the top throughout.

Then they said to one another, “Let’s not tear it, but cast lots for it to decide whose it will be,” that the Scripture might be fulfilled, which says, “They parted my garments among them. For my cloak they cast lots.” Therefore the soldiers did these things.

Those who passed by blasphemed him, wagging their heads,

and saying, “You who destroy the temple, and build it in three days, save yourself! If you are the Son of God, come down from the cross!”

Likewise the chief priests also mocking, with the scribes, the Pharisees, and the elders, said,



\chapterornament
\section*{Page 54}

\lettrine{“}{He saved others, but he can’t save himself. If he is the King of Israel, let him come down from the cross now, and we will believe in him.}

He trusts in God. Let God deliver him now, if he wants him; for he said, ‘I am the Son of God.’ ”

One of the criminals who was hanged insulted him, saying, “If you are the Christ, save yourself and us!”

But the other answered, and rebuking him said, “Don’t you even fear God, seeing you are under the same condemnation?

And we indeed justly, for we receive the due reward for our deeds, but this man has done nothing wrong.”

Jesus said, “Father, forgive them, for they don’t know what they are doing.” Dividing his garments among them, they cast lots.

But standing by Jesus’ cross were his mother, his mother’s sister, Mary the wife of Clopas, and Mary Magdalene.

Therefore when Jesus saw his mother, and the disciple whom he loved standing there, he said to his mother, “Woman, behold, your son!”

Then he said to the disciple, “Behold, your mother!” From that hour, the disciple took her to his own home.

About the ninth hour Jesus cried with a loud voice, saying, “Eli, Eli, lima sabachthani?”  That is, “My God, my God, why have you forsaken me?”

Some of them who stood there, when they heard it, said, “This man is calling Elijah.”

Immediately one of them ran, and took a sponge, and filled it with vinegar, and put it on a reed, and gave him a drink.

The rest said, “Let him be. Let’s see whether Elijah comes to save him.”

Jesus cried again with a loud voice, and yielded up his spirit.

Many women were there watching from afar, who had followed Jesus from Galilee, serving him.

Among them were Mary Magdalene, Mary the mother of James and Joses, and the mother of the sons of Zebedee.

Therefore the Jews, because it was the Preparation Day, so that the bodies wouldn’t remain on the cross on the Sabbath (for that Sabbath was a special one), asked of Pilate that their legs might be broken, and that they might be taken away.

Therefore the soldiers came, and broke the legs of the first, and of the other who was crucified with him;

but when they came to Jesus, and saw that he was already dead, they didn’t break his legs.

However one of the soldiers pierced his side with a spear, and immediately blood and water came out.



\chapterornament
\section*{Page 55}

\lettrine{A}{fter these things, Joseph of Arimathaea, being a disciple of Jesus, but secretly for fear of the Jews, asked of Pilate that he might take away Jesus’ body. Pilate gave him permission. He came therefore and took away his body.}

Nicodemus, who at first came to Jesus by night, also came bringing a mixture of myrrh and aloes, about a hundred Roman pounds.

So they took Jesus’ body, and bound it in linen cloths with the spices, as the custom of the Jews is to bury.

Now in the place where he was crucified there was a garden. In the garden was a new tomb in which no man had ever yet been laid.

Then because of the Jews’ Preparation Day (for the tomb was near at hand) they laid Jesus there.

and laid it in his own new tomb, which he had cut out in the rock, and he rolled a great stone against the door of the tomb, and departed.
